### 2022-12-18 submission.tex
first commit

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage[margin=2cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usepackage[export]{adjustbox}
\usepackage{indentfirst}
\usepackage{wrapfig}
\usepackage{listings}
\usepackage{color}
\usepackage{enumerate}
\usepackage{amssymb, bm}
\usepackage{csvsimple}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usepackage{lscape}
\newcommand\tab[1][1cm]{\hspace*{#1}}
\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}
\lstset{ %
    backgroundcolor=\color{gray!10!white},
  basicstyle=\tiny, %footnotesize,        % the size of the fonts that are used for the code
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{mygreen},    % comment style
  deletekeywords={...},            % if you want to delete keywords from the given language
  escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
  frame=single,	                   % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\color{blue},       % keyword style
  language=Python,                 % the language of the code
  morekeywords={*,...},           % if you want to add more keywords to the set
  numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=10pt,                   % how far the line-numbers are from the code
  numberstyle=\tiny\color{mygray}, % the style that is used for the line-numbers
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
  showstringspaces=false,          % underline spaces within strings only
  showtabs=false,                  % show tabs within strings adding particular underscores
  stepnumber=1,                    % the step between two line-numbers. If it's 1, each line will be numbered
  stringstyle=\color{mymauve},     % string literal style
  tabsize=1,	                   % sets default tabsize to 2 spaces
  title=\lstname                   % show the filename of files included with \lstinputlisting; also try caption instead of title
}
\graphicspath{ }
\usetikzlibrary{arrows}

\title{\textbf{COMP0083 Convex Optimisation Assignment}}
%\author{Jian Shu (James) Wu \\ }
\date{Jan 9, 2023}

\begin{document}
\maketitle
\section*{Part 1}
\subsection*{1.1}
\subsubsection*{1.1.1}

\end{document}

### 2022-12-25 submission.tex
add some written answers and classes for algorithms

@@
 \usepackage{lscape}
+\usepackage{amsmath}
 \newcommand\tab[1][1cm]{\hspace*{#1}}
@@
 \subsection*{1.1}
-\subsubsection*{1.1.1}
+
+The function for (a):
+
+\[\max\{ax+b, x^4-5, e^{x^2}\}\]
+
+is a convex function.
+
+\subsection*{1.2}
+
+For the function:
+
+\[f(x) = \begin{cases}
+      -x \tab \text{if } x \in ]-1, 0] \\
+      x^2 \tab \text{if } x \geq 0 \\
+   \end{cases}\]
+
+the sub-differential is:
+
+\[\partial f(x) = \begin{cases}
+      -1 \tab \text{if } x \in ]-1, 0[ \\
+      \left[0, 1\right] \tab \text{if } x = 0  \\
+      2x \tab \text{if } x > 0 \\
+   \end{cases}\]
+
+corresponding to Figure (a).
+
+\subsection*{1.3}
+For a function:
+\[f(x) = \langle Ax, x \rangle + \langle x, b\rangle + c\]
+
+where A is a square matrix not necessarily symmetric, the gradient is (a):
+
+\[\nabla f(x) = A^*x + Ax + b\]
+
+\subsection*{1.4}
+
+The Fenchel conjugate of $f(x) = g(2x)$ is (a):
+
+\[f^*(u) = g^*(u/2)\]
+
+\subsection*{1.5}
+
+The solution to the dual problem is (c):
+
+\[\bar{u} = (\textbf{K}+\lambda n \textbf{Id})^{-1} y\]
+
+\section*{Part 2}
+\subsection*{2.1}
+
+\subsection*{2.1.1}
+
+Given :
+\[ f(x) = \begin{cases}
+      +\infty \tab \text{if } x \leq 0 \\
+      -\log x \tab \text{if } x > 0 \\
+   \end{cases}\]
+
+The Fenchel conjugate is defined:
+
+\[f^*(u) = \sup_{x\in \mathcal{X}} \{ \langle x, u\rangle +  \log x \}\]
+
+To find the supremum, we can take the partial derivative with respect to $x$, set to zero, and solve for x:
+
+\[\frac{\partial}{\partial x}(ux + logx) = u + \frac{1}{x} = 0\]
+
+Thus, the supremum above is solved when $x=\frac{-1}{u}$:
+
+\[f^*(u) = u \left(\frac{-1}{u} \right) +  \log\left(\frac{-1}{u} \right) \]
+
+Simplifying, we have the Fenchel conjugate:
+
+\[f^*(u) = - (1 + \log u)\]
+
+\subsection*{2.1.2} Given:
+
+\[f(x) = x^2\]
+
+The Fenchel conjugate is defined:
+
+\[f^*(u) = \sup_{x\in \mathcal{X}} \{ \langle x, u\rangle - x^2 \}\]
+
+We can compute the partial derivative:
+
+\[\frac{\partial}{\partial x}(ux - x^2) = u - 2x = 0\]
+
+Thus, the supremum is solved when $x=\frac{u}{2}$:
+
+\[f^*(u) = u \left(\frac{u}{2} \right) -  \left(\frac{u}{2} \right)^2 \]
+
+Simplifying, we have the Fenchel conjugate:
+
+\[f^*(u) =  \frac{u^2}{4}\]
+
+\subsection*{2.1.3}Given:
+
+\[f(x) = i_{\left[0, 1\right]}\]
+
+The Fenchel conjugate is defined:
+
+\[f^*(u) = \sup_{x\in \mathcal{X}} \{ \langle x, u\rangle - i_{\left[0, 1\right]} \}\]
+
+Thus,
+
+\[f^*(u) = \sup_{x\in \left[0, 1\right]} \{ \langle x, u\rangle \}\]
+
+We can see the Fenchel conjugate is:
+
+%\[ f^*(u) = \begin{cases}
+%      0 \tab \text{if } u < 0 \\
+%      u \tab \text{if } u \geq 0 \\
+%   \end{cases}\]
+
+\[ f^*(u) = \max(0, u)}\]
+
+\subsection*{2.2.1}
+
+Given $f$ a proper convex function, to prove by induction Jensen's inequality:
+
+\[f\left(\sum_{i=1}^n \lambda_i x_i\right) \leq \sum_{i=1}^{n} \lambda_i f(x_i)\]
+
+for all $x_1, ... x_n \in \mathcal{X}$ and for all $\lambda_1, ..., \lambda_n \in \mathbb{R}_+$ with $\sum_{i=1}^n \lambda_i = 1$, we start with the definition of convexity which states:
+
+\[f(\lambda_1 x_1 + \lambda_2 x_2) \leq \lambda_1 f(x_1) + \lambda_2 f(x_2)\]
+
+for all $x1, x2 \in \mathcal{X}$ and $\lambda_1, \lambda_2 \in \mathbb{R}_+$ with $\lambda_1 + \lambda_2 = 1$. In this base case $n=2$.
+
+Our inductive step will prove that the inequality continues to hold for $n+1$:
+
+\[f\left(\sum_{i=1}^{n+1} \lambda_i x_i\right) = f\left(\sum_{i=1}^n \lambda_i x_i + \lambda_{n+1} x_{n+1}\right)\]
+
+We can insert the term $(1-\lambda_{n+1})$:
+
+\[f\left(\sum_{i=1}^{n+1} \lambda_i x_i\right) = f\left((1-\lambda_{n+1}) \sum_{i=1}^n \frac{\lambda_i}{1-\lambda_{n+1}} x_i + \lambda_{n+1} x_{n+1}\right)\]
+
+If we define $\bar{x} = \sum_{i=1}^n \frac{\lambda_i}{1-\lambda_{n+1}} x_i$ and we are back to our $n=2$ base case, so we know from convexity:
+
+\[ f\left((1-\lambda_{n+1})\bar{x} + \lambda_{n+1} x_{n+1}\right) \leq (1-\lambda_{n+1}) f\left( \bar{x}\right) + \lambda_{n+1} f\left( x_{n+1}\right)\]
+
+Rewriting this,
+
+\[f\left(\sum_{i=1}^{n+1} \lambda_i x_i\right) \leq (1-\lambda_{n+1}) f\left( \sum_{i=1}^n \frac{\lambda_i}{1-\lambda_{n+1}} x_i\right) + \lambda_{n+1} f\left( x_{n+1}\right)\]
+
+We know that the first term on the right hand side:
+
+\[f\left( \sum_{i=1}^n \frac{\lambda_i}{1-\lambda_{n+1}} x_i\right) = \frac{1}{{1-\lambda_{n+1}}}f\left( \sum_{i=1}^n \lambda_i x_i\right) \leq \frac{1}{{1-\lambda_{n+1}}} \sum_{i=1}^n \lambda_i f \left(x_i\right) =  \sum_{i=1}^n \frac{\lambda_i}{{1-\lambda_{n+1}}}  f \left(x_i\right)\]
+
+Thus can upper bound our previous inequality:
+
+\[f\left(\sum_{i=1}^{n+1} \lambda_i x_i\right)  \leq \sum_{i=1}^n \frac{\lambda_i}{{1-\lambda_{n+1}}}  f \left(x_i\right) + \lambda_{n+1} f\left( x_{n+1}\right) =\sum_{i=1}^{n+1} \frac{\lambda_i}{{1-\lambda_{n+1}}} f \left(x_i\right) \]
+
+proving our inductive step and Jensen's inequality as required. \square
+
+\subsection*{2.2.2}
+The characterisation of convexity states:
+
+\[\text{f is convex} \leftrightarrow \langle \nabla f(x) - \nabla f(y), x-y \rangle \geq 0\]
+
+$\forall x, y \in dom f$.
+
+For $f(x) = -\log(x)$, $\nabla f(x) = \frac{-1}{x}$:
+
+\[\langle \nabla f(x) - \nabla f(y), x-y \rangle = \langle \frac{1}{y}-\frac{1}{x}, x-y \rangle\]
+
+If $x>y$ we have the following:
+
+\begin{gather*}
+    \frac{1}{y} > \frac{1}{x}\\
+    x-y>0\\
+    \frac{1}{y} - \frac{1}{x} > 0\\
+\end{gather*}
+
+and
+\[\langle \frac{1}{y}-\frac{1}{x}, x-y \rangle > 0\]
+
+If $x<y$ we have the following:
+\begin{gather*}
+    \frac{1}{y} < \frac{1}{x}\\
+    x-y<0\\
+    \frac{1}{y} - \frac{1}{x} < 0\\
+\end{gather*}
+
+and
+\[\langle \frac{1}{y}-\frac{1}{x}, x-y \rangle > 0\]
+
+If $x=y$ we have the following:
+\begin{gather*}
+    \frac{1}{y} = \frac{1}{x}\\
+    x-y=0\\
+    \frac{1}{y} - \frac{1}{x} = 0\\
+\end{gather*}
+
+and
+\[\langle \frac{1}{y}-\frac{1}{x}, x-y \rangle = 0\]
+
+Thus, $\forall x, y \in dom f, \langle \nabla f(x) - \nabla f(y), x-y \rangle \geq 0$ so $f(x) = -\log(x)$ is convex. \square
+
+\subsection*{2.2.3}
+
+Given Jensen's inequality in 2.2.1 for a convex function $f(x)$:
+
+\[f\bigg(\sum_{i=1}^n \lambda_i x_i\bigg) \leq \sum_{i=1}^n \lambda_i f(x_i)\]
+
+and having proved in 2.2.2 that $f(x) = -\log(x)$ is convex, we can write:
+
+\[-\log\bigg(\sum_{i=1}^n \lambda_i x_i\bigg) \leq - \sum_{i=1}^n \lambda_i \log(x_i)\]
+
+Rearranging:
+
+\[\sum_{i=1}^n \lambda_i x_i \geq \exp \bigg(\sum_{i=1}^n \log(x_i^{\lambda_i}) \bigg)\]
+
+Choosing $\lambda_i = \frac{1}{n}$:
+
+\[\frac{1}{n}\sum_{i=1}^n x_i \geq \exp \bigg(\sum_{i=1}^n \log(x_i^{\frac{1}{n}}) \bigg)\]
+
+The sum of logarithms is the logarithm of the products:
+
+\[\frac{1}{n}\sum_{i=1}^n x_i \geq \exp \bigg(\log((x_1 \cdots x_n )^{\frac{1}{n}})\bigg)\]
+
+Thus we get our inequality:
+
+\[\frac{1}{n}\sum_{i=1}^n x_i \geq \sqrt[n]{x_1 \cdots x_n}\] \square
+\subsection*{2.3}
+\subsection*{2.4}
+\subsection*{2.5}
+  Conditions for the existence of minimizers: $f$ is closed and coercive.
+
+  Conditions for the uniqueness of minimizers: $f$ is strictly convex.
+\subsection*{2.6}
+
+\newpage
+\section*{Part 3}
+
+\newpage
+\section*{Part 4}
+
+We are given the primal problem:
+
+\[\min_{w \in \mathcal{H}} \frac{\lambda}{n} \sum_{i=1}^{n} (1-y_i \langle w, \Lambda(x_i)\rangle)_{+} + \frac{\lambda}{2}\| w\|^2\]
+
+We can express in the form:
+
+\[\min_{w \in \mathcal{H}} g(w) + f(w)\]
+
+where:
+
+\[g(w) = \frac{\lambda}{n} \sum_{i=1}^{n} (1-y_i \langle w, \Lambda(x_i)\rangle)_{+}\]
+
+and
+
+\[f(w) = \frac{\lambda}{2}\| w\|^2\]
+
+The corresponding dual problem:
+
+\[\min_{\alpha \in \mathbb{R}^n} \frac{1}{2} \langle \mathbf{K}_y \alpha, \alpha \rangle - \langle \textbf{1}_n, \alpha\rangle + \sum_{i=1}^{n} i_{\left[0, \frac{1}{\lambda n}\right]}(\alpha_i)\]
+
+We can express in the form:
+
+\[\min_{\alpha \in \mathbb{R}^n} g^*(\alpha) + f^*(\alpha)\]
+
+\subsection*{4.1}
 

### 2022-12-25 gradient_algorithms.py

@@ -1,0 +1,109 @@
+from problems import Problem
+import numpy as np
+from abc import ABC, abstractmethod
+
+
+class GradientAlgorithm(ABC):
+    def __init__(self, problem: Problem):
+        self.problem = problem
+
+    @abstractmethod
+    def run(self, x, lambda_parameter, number_of_steps):
+        pass
+
+
+class ProximalStochasticGradientAlgorithm(GradientAlgorithm):
+    def __init__(self, problem: Problem):
+        super().__init__(problem)
+
+    def generate_random_data_index(self):
+        return np.random.choice(self.problem.n)
+
+    def calculate_gamma(self, k):
+        """
+        A matrix (number of points, number of dimensions)
+        k: step number
+        """
+        return self.problem.n / (
+            np.linalg.norm(self.problem.a_matrix) ** 2 * np.sqrt(k + 1)
+        )
+
+    def run(self, x, lambda_parameter, number_of_steps):
+        for k in range(number_of_steps):
+            i = self.generate_random_data_index()
+            gamma_k = self.calculate_gamma(k)
+            x = self.problem.proximity_operator(
+                x=x
+                - gamma_k
+                * (self.problem.a_matrix_i(i) @ x - self.problem.y[i])
+                * self.problem.a_matrix_i(i),  # (number of dimensions, )
+                gamma=gamma_k,
+                lambda_parameter=lambda_parameter,
+            )
+        return x
+
+
+class RandomizedCoordinateProximalGradientAlgorithm(GradientAlgorithm):
+    def __init__(self, problem: Problem):
+        super().__init__(problem)
+
+    def generate_random_dimension_index(self):
+        return np.random.choice(self.problem.d)
+
+    def calculate_gamma(self, j):
+        """
+        A matrix (number of points, number of dimensions)
+        j: dimension
+        """
+        return self.problem.n / (np.linalg.norm(self.problem.a_matrix_j(j)) ** 2)
+
+    def run(self, x, lambda_parameter, number_of_steps):
+        for k in range(number_of_steps):
+            j = self.generate_random_dimension_index()
+            gamma_j = self.calculate_gamma(j)
+            x[j] = self.problem.proximity_operator(
+                x=x[j]
+                - (gamma_j / self.problem.n)
+                * self.problem.a_matrix_j(j)
+                @ (
+                    self.problem.a_matrix @ x - self.problem.y
[…]

### 2022-12-25 problems.py

@@ -1,0 +1,157 @@
+from __future__ import annotations
+import random
+from abc import ABC, abstractmethod
+import numpy as np
+from sklearn.datasets import make_moons
+
+
+class Problem(ABC):
+    def __init__(self, a_matrix: np.ndarray, y: np.ndarray):
+        """
+
+        :param a_matrix: design matrix (number of points, number of dimensions)
+        :param y: response vector (number of points, 1)
+        """
+        self.a_matrix = a_matrix
+        self.y = y
+
+    @property
+    def n(self):
+        return self.a_matrix.shape[0]
+
+    @property
+    def d(self):
+        return self.a_matrix.shape[1]
+
+    def a_matrix_i(self, i):
+        return self.a_matrix[i, :]
+
+    def a_matrix_j(self, j):
+        return self.a_matrix[:, j]
+
+    @staticmethod
+    @abstractmethod
+    def proximity_operator(x, gamma, lambda_parameter):
+        pass
+
+    @abstractmethod
+    def loss(self, x, lambda_parameter: float):
+        pass
+
+    @staticmethod
+    @abstractmethod
+    def generate(*args, **kwargs) -> Problem:
+        pass
+
+
+class SparseProblem(Problem):
+    def __init__(
+        self,
+        a_matrix: np.ndarray,
+        y: np.ndarray,
+        x_sparse: np.ndarray,
+        sparsity: int,
+        std: float,
+    ):
+        self.x_sparse = x_sparse
+        self.sparsity = sparsity
+        self.std = std
+        super().__init__(a_matrix, y)
+
+    @staticmethod
+    def proximity_operator(x, gamma, lambda_parameter):
+        rho = gamma * lambda_parameter
+        return np.sign(x - rho) * np.maximum(0, np.abs(x) - rho)
+
+    def loss(self, x, lambda_parameter: float):
+        return (1 / (2 * self.n)) * np.linalg.norm(
+            self.a_matrix @ x - self.y
+        ) ** 2 + lambda_parameter * np.linalg.norm(x, ord=1)
+
+    @staticmethod
+    def generate(
+        number_of_samples, number_of_dimensions, sparsity: int, std=0.06
+    ) -> SparseProblem:
+        """
+        Generate xs vectors with entries in [0.5, 1] and [-1, -0.5] respectively.
+
+        :param number_of_samples:
+        :param number_of_dimensions:
+        :param sparsity:
[…]

### 2022-12-27 gradient_algorithms.py
abstract gradient calculation in algorithms and add specific grad for each problem definition

@@ -1,6 +1,8 @@
-from problems import Problem
-import numpy as np
 from abc import ABC, abstractmethod
+
+import numpy as np
+
+from problems import Problem
 
 
 class GradientAlgorithm(ABC):
@@ -13,7 +15,8 @@ class GradientAlgorithm(ABC):
 
 
 class ProximalStochasticGradientAlgorithm(GradientAlgorithm):
-    def __init__(self, problem: Problem):
+    def __init__(self, problem: Problem, is_ergodic_mean: bool):
+        self.is_ergodic_mean = is_ergodic_mean
         super().__init__(problem)
 
     def generate_random_data_index(self):
@@ -28,7 +31,25 @@ class ProximalStochasticGradientAlgorithm(GradientAlgorithm):
             np.linalg.norm(self.problem.a_matrix) ** 2 * np.sqrt(k + 1)
         )
 
-    def run(self, x, lambda_parameter, number_of_steps):
+    def _run_ergodic_mean(self, x, lambda_parameter, number_of_steps):
+        gamma_normaliser = 0
+        x_bar_unnormalised = np.zeros(x.shape)
+        for k in range(number_of_steps):
+            i = self.generate_random_data_index()
+            gamma_k = self.calculate_gamma(k)
+            x = self.problem.proximity_operator(
+                x=x
+                - gamma_k
+                * (self.problem.a_matrix_i(i) @ x - self.problem.y[i])
+                * self.problem.a_matrix_i(i),  # (number of dimensions, )
+                gamma=gamma_k,
+                lambda_parameter=lambda_parameter,
+            )
+            gamma_normaliser += gamma_k
+            x_bar_unnormalised += gamma_k * x
+        return x_bar_unnormalised / gamma_normaliser
+
+    def _run_not_ergodic_mean(self, x, lambda_parameter, number_of_steps):
         for k in range(number_of_steps):
             i = self.generate_random_data_index()
             gamma_k = self.calculate_gamma(k)
@@ -41,6 +62,12 @@ class ProximalStochasticGradientAlgorithm(GradientAlgorithm):
                 lambda_parameter=lambda_parameter,
             )
         return x
+
+    def run(self, x, lambda_parameter, number_of_steps):
+        if self.is_ergodic_mean:
+            return self._run_ergodic_mean(x, lambda_parameter, number_of_steps)
+        else:
+            return self._run_not_ergodic_mean(x, lambda_parameter, number_of_steps)
 
 
[…]

### 2022-12-27 problems.py

@@ -1,6 +1,8 @@
 from __future__ import annotations
+
 import random
 from abc import ABC, abstractmethod
+
 import numpy as np
 from sklearn.datasets import make_moons
 
@@ -43,6 +45,14 @@ class Problem(ABC):
     def generate(*args, **kwargs) -> Problem:
         pass
 
+    @abstractmethod
+    def grad_f_j(self, x: np.ndarray, j: int) -> float:
+        pass
+
+    @abstractmethod
+    def grad_f(self, x: np.ndarray) -> float:
+        return self.a_matrix @ x
+
 
 class SparseProblem(Problem):
     def __init__(
@@ -67,6 +77,12 @@ class SparseProblem(Problem):
         return (1 / (2 * self.n)) * np.linalg.norm(
             self.a_matrix @ x - self.y
         ) ** 2 + lambda_parameter * np.linalg.norm(x, ord=1)
+
+    def grad_f_j(self, x: np.ndarray, j) -> float:
+        return self.a_matrix_j(j) @ (self.a_matrix @ x - self.y)
+
+    def grad_f(self, x: np.ndarray) -> float:
+        pass
 
     @staticmethod
     def generate(
@@ -102,7 +118,7 @@ class SparseProblem(Problem):
         )
 
 
-class SVMProblem(Problem):
+class HalfMoonsProblem(Problem):
     def __init__(self, x: np.ndarray, y: np.ndarray, sigma: float):
         self.x = x
         self.sigma = sigma
@@ -140,18 +156,35 @@ class SVMProblem(Problem):
         return out
 
     def loss(self, x, lambda_parameter: float):
-        pass
+        indicator = (
+            np.float("inf")
+            if np.sum(np.logical_or((x < 0), (x > (1 / (lambda_parameter * self.n)))))
+            else 0
+        )
+        return 0.5 * x.T @ self.a_matrix @ x - np.sum(x) + indicator
+
+    def grad_f_j(self, x: np.ndarray, j) -> float:
+        return self.a_matrix_j(j) @ x
+
+    def grad_f(self, x: np.ndarray) -> float:
+        return self.a_matrix @ x
 
     @staticmethod
     def generate(
         number_of_samples: int, noise: float, sigma: float, random_state: int
-    ) -> SVMProblem:
+    ) -> HalfMoonsProblem:
         x, y = make_moons(
             n_samples=number_of_samples, noise=noise, random_state=random_state
         )
         y = 2 * y - 1
[…]

### 2023-01-04 submission.tex
update report

@@
 
-%\[ f^*(u) = \begin{cases}
-%      0 \tab \text{if } u < 0 \\
-%      u \tab \text{if } u \geq 0 \\
-%   \end{cases}\]
-
 \[ f^*(u) = \max(0, u)}\]
@@
 \subsection*{2.3}
+
+Given a polytope $C=co(a_1, ..., a_m)$ in $X$, we know that any point $x \in C$ can be expressed:
+
+\[x = \sum_{i=1}^{m} \lambda_i a_i \]
+
+where $\sum_{i=1}^{m} \lambda_i = 1$. Knowing that $f$ is a convex function on $C$, we can write Jensen's inequality for :
+
+\[f\left(\sum_{i=1}^{m} \lambda_i a_i \right) \leq \sum_{i=1}^{m} \lambda_i f(a_i)\]
+
+where $\sum_{i=1}^{m} \lambda_i = 1$.
+
+Because $\sum_{i=1}^{m} \lambda_i f(a_i)$ is a weighted average of $f(a_i)$'s we know that:
+
+\[\sum_{i=1}^{m} \lambda_i f(a_i) \leq \max_{\{a_i\}_{i=1}^m} f(a_i)\]
+
+the weighted average will always be less than or equal to the maximum $f(a_i)$ value. Thus, we know that:
+
+\[f\left(\sum_{i=1}^{m} \lambda_i a_i \right) \leq \max_{\{a_i\}_{i=1}^m} f(a_i)\]
+
+So the maximum of the convex function $f$ on $C$ is attained at one of the vertices $a_1, ..., a_m$.
+\square
+
 \subsection*{2.4}
+
+To prove that the function $f(x, y) = \| x-2y\|^2$ is convex, we will show that the Hessian is a positive semi-definite matrix. The Hessian is defined as:
+
+\[
+\nabla^2 f(x, y) = \begin{bmatrix}
+                 \frac{\partial f(x, y)}{\partial^2 x} & \frac{\partial f(x, y)}{\partial y \partial x}\\
+                 \frac{\partial f(x, y)}{\partial x \partial y} & \frac{\partial f(x, y)}{\partial^2 y}
+         \end{bmatrix}
+\]
+
+Calculating each term:
+
+\[\nabla^2 f(x, y) = \begin{bmatrix}
+                 2 & 4\\
+                 4 & 8
+         \end{bmatrix}
+\]
+
+we see that the Hessian is positive semi-definite and so the function is jointly convex.
+
 \subsection*{2.5}
-  Conditions for the existence of minimizers: $f$ is closed and coercive.
+  The conditions for the existence of minimizers is that $f$ is closed and coercive.
+  The conditions for the uniqueness of minimizers is that $f$ is strictly convex.
+Thus for the existence and uniqueness of minimizers for a convex function $f$, we require that $f$ is closed, coercive, and strictly convex.
 
-  Conditions for the uniqueness of minimizers: $f$ is strictly convex.
 \subsection*{2.6}
@@
 \section*{Part 3}
+
+Our problem:
+
+\[\min_{\textbf{x} \in \mathbb{R}^d} \frac{1}{2n} \|\textbf{Ax}-\textbf{y}\|^2 + \lambda \|x\|_1\]
+
+where $\textbf{A} \in \mathbb{R}^{n\times d}$, $\textbf{x} \in \mathbb{R}^{d \times 1}$, $\textbf{y} \in \mathbb{R}^{n \times 1}$, $n$ is the number of data points, and $d$ is the number of dimensions.
+
+Equivalently, our problem:
+
+\[\min_{\textbf{x} \in \mathbb{R}^d} \frac{1}{2n} \sum_{i=1}^n \left(\left\langle \textbf{a}^i, \textbf{x} \right\rangle - y_i \right)^2 + \lambda \|x\|_1\]
+
+where $\textbf{a}^i \in \mathbb{R}^{1 \times d}$ is the $i^{th}$ row of \textbf{A}, $i=1, ..., n$.
+
+\subsection*{3.1}
+The Proximal Stochastic Gradient Algorithm:
+
+\[\textbf{x}^{k+1} = prox_{\gamma_k \lambda \|\cdot\|_1} \left(\textbf{x}^k - \gamma_k \left(\left\langle \textbf{a}^{i_k}, \textbf{x}^{k}\right\rangle - y_{i_k}\right)\textbf{a}^{i_k}\right)\]
+
+where:
+
+\[\gamma_k = \frac{n}{\|A\|^2 \sqrt{k+1}}\]
+
+and
+
+\[prox_{\gamma_k \lambda \|\cdot\|_1}(x) = soft_{\gamma_k \lambda}(x) = \begin{cases}
+      0, &  \text{if } |x| \leq \gamma_k \lambda \\
+      x-\gamma_k \lambda, & \text{if } x > \gamma_k \lambda \\
+      x+\gamma_k \lambda, & \text{if } x < - \gamma_k \lambda \\
+   \end{cases}\]
+
+and $i_k$ is sampled uniformly from $\{1, ..., n\}$ at each step $k$.
+
+\subsection*{3.2}
+
+The Randomized Coordinate Proximal Gradient Algorithm:
+
+\[x_j^{k+1} = \begin{cases}
+      soft_{\gamma_j \lambda}(x_j^k - \frac{\gamma_j}{n} \left\langle a_j, \textbf{A} \textbf{x}^k - y\right\rangle), & \text{if } j = j_k \\
+      x_j^k, & \text{otherwise} \\
+   \end{cases}
+\]
+
+where we define $\textbf{a}_j \in \mathbb{R}^{n \times 1}$ as the $j^{th}$ column of $\textbf{A}$, $i=1, ..., d$.
+
+\[\gamma_j = \frac{n}{\|\textbf{a}_j\|^2}\]
+
+and $j_k$ is sampled uniformly from $\{1, ..., d\}$ at each step $k$.
 

### 2023-01-04 constants.py
add figures for part 3

@@ -1,0 +1,3 @@
+OUTPUTS_FOLDER = "outputs"
+
+DEFAULT_SEED = 0

### 2023-01-04 gradient_algorithms.py

@@ -1,125 +1,0 @@
-from abc import ABC, abstractmethod
-
-import numpy as np
-
-from problems import Problem
-
-
-class GradientAlgorithm(ABC):
-    def __init__(self, problem: Problem):
-        self.problem = problem
-
-    @abstractmethod
-    def run(self, x, lambda_parameter, number_of_steps):
-        pass
-
-
-class ProximalStochasticGradientAlgorithm(GradientAlgorithm):
-    def __init__(self, problem: Problem, is_ergodic_mean: bool):
-        self.is_ergodic_mean = is_ergodic_mean
-        super().__init__(problem)
-
-    def generate_random_data_index(self):
-        return np.random.choice(self.problem.n)
-
-    def calculate_gamma(self, k):
-        """
-        A matrix (number of points, number of dimensions)
-        k: step number
-        """
-        return self.problem.n / (
-            np.linalg.norm(self.problem.a_matrix) ** 2 * np.sqrt(k + 1)
-        )
-
-    def _run_ergodic_mean(self, x, lambda_parameter, number_of_steps):
-        gamma_normaliser = 0
-        x_bar_unnormalised = np.zeros(x.shape)
-        for k in range(number_of_steps):
-            i = self.generate_random_data_index()
-            gamma_k = self.calculate_gamma(k)
-            x = self.problem.proximity_operator(
-                x=x
-                - gamma_k
-                * (self.problem.a_matrix_i(i) @ x - self.problem.y[i])
-                * self.problem.a_matrix_i(i),  # (number of dimensions, )
-                gamma=gamma_k,
-                lambda_parameter=lambda_parameter,
-            )
-            gamma_normaliser += gamma_k
-            x_bar_unnormalised += gamma_k * x
-        return x_bar_unnormalised / gamma_normaliser
-
-    def _run_not_ergodic_mean(self, x, lambda_parameter, number_of_steps):
-        for k in range(number_of_steps):
-            i = self.generate_random_data_index()
-            gamma_k = self.calculate_gamma(k)
-            x = self.problem.proximity_operator(
-                x=x
-                - gamma_k
-                * (self.problem.a_matrix_i(i) @ x - self.problem.y[i])
-                * self.problem.a_matrix_i(i),  # (number of dimensions, )
-                gamma=gamma_k,
-                lambda_parameter=lambda_parameter,
-            )
-        return x
-
-    def run(self, x, lambda_parameter, number_of_steps):
[…]

### 2023-01-04 main.py

@@ -1,0 +1,3 @@
+from part_3 import part_3
+
+part_3()

### 2023-01-04 part_3.py

@@ -1,0 +1,151 @@
+import os
+from typing import List
+
+import matplotlib.pyplot as plt
+import numpy as np
+
+from constants import DEFAULT_SEED, OUTPUTS_FOLDER
+from src.gradient_algorithms import (
+    ProximalStochasticGradientAlgorithm,
+    RandomizedCoordinateProximalGradientAlgorithm,
+)
+from src.problems import SparseProblem
+
+
+def plot_loss(
+    losses: List[List[float]], labels: List[str], algorithm: str, save_path: str
+):
+    fig = plt.figure()
+    fig.set_figwidth(10)
+    fig.set_figheight(5)
+    for i, loss in enumerate(losses):
+        plt.plot(loss, label=labels[i])
+    if len(labels) > 1:
+        plt.legend()
+    plt.xlabel("Iteration")
+    plt.ylabel("Loss")
+    plt.title(f"Loss ({algorithm})")
+    plt.savefig(save_path, bbox_inches="tight")
+    plt.close()
+
+
+def plot_dimensions(
+    x: np.ndarray, x_sparse: np.ndarray, algorithm: str, save_path: str
+) -> None:
+    """
+    Plot dimensions of solution and actual sparse vector
+
+    :param x: solution (number of dimension, 1)
+    :param x_sparse: actual sparse vector (number of dimensions, 1)
+    :param algorithm: algorithm generating the solution
+    :param save_path: path to save figure
+    """
+    x_sparse = x_sparse.reshape(-1)
+    x = x.reshape(-1)
+    d = len(x)
+
+    fig, ax = plt.subplots(1, 1)
+    fig.set_figwidth(10)
+    fig.set_figheight(5)
+    # True variable x_sparse
+    ax.stem(
+        np.arange(d)[np.abs(x_sparse) > 0],
+        x_sparse[np.abs(x_sparse) > 0],
+        label="$x^*$",
+    )
+    ax.stem(
+        np.arange(d)[np.abs(x) > 0],
+        x[np.abs(x) > 0],
+        label=r"$x_{\tau, \lambda}$",
+        linefmt="k:",
+        markerfmt="k^",
+    )
+    ax.axhline(0.0, color="red")
+    ax.set_xlim([-10 + 0, d + 10])
+    ax.set_ylim([-1.1, 1.1])
+    ax.set_xlabel("index $i$")
+    ax.set_ylabel("$x_i$")
+    ax.legend()
+    plt.title(f"Sparse Solution vs Actual ({algorithm})")
[…]

### 2023-01-04 problems.py

@@ -1,190 +1,0 @@
-from __future__ import annotations
-
-import random
-from abc import ABC, abstractmethod
-
-import numpy as np
-from sklearn.datasets import make_moons
-
-
-class Problem(ABC):
-    def __init__(self, a_matrix: np.ndarray, y: np.ndarray):
-        """
-
-        :param a_matrix: design matrix (number of points, number of dimensions)
-        :param y: response vector (number of points, 1)
-        """
-        self.a_matrix = a_matrix
-        self.y = y
-
-    @property
-    def n(self):
-        return self.a_matrix.shape[0]
-
-    @property
-    def d(self):
-        return self.a_matrix.shape[1]
-
-    def a_matrix_i(self, i):
-        return self.a_matrix[i, :]
-
-    def a_matrix_j(self, j):
-        return self.a_matrix[:, j]
-
-    @staticmethod
-    @abstractmethod
-    def proximity_operator(x, gamma, lambda_parameter):
-        pass
-
-    @abstractmethod
-    def loss(self, x, lambda_parameter: float):
-        pass
-
-    @staticmethod
-    @abstractmethod
-    def generate(*args, **kwargs) -> Problem:
-        pass
-
-    @abstractmethod
-    def grad_f_j(self, x: np.ndarray, j: int) -> float:
-        pass
-
-    @abstractmethod
-    def grad_f(self, x: np.ndarray) -> float:
-        return self.a_matrix @ x
-
-
-class SparseProblem(Problem):
-    def __init__(
-        self,
-        a_matrix: np.ndarray,
-        y: np.ndarray,
-        x_sparse: np.ndarray,
-        sparsity: int,
-        std: float,
-    ):
-        self.x_sparse = x_sparse
-        self.sparsity = sparsity
-        self.std = std
-        super().__init__(a_matrix, y)
-
-    @staticmethod
-    def proximity_operator(x, gamma, lambda_parameter):
-        rho = gamma * lambda_parameter
-        return np.sign(x - rho) * np.maximum(0, np.abs(x) - rho)
-
-    def loss(self, x, lambda_parameter: float):
-        return (1 / (2 * self.n)) * np.linalg.norm(
-            self.a_matrix @ x - self.y
-        ) ** 2 + lambda_parameter * np.linalg.norm(x, ord=1)
-
-    def grad_f_j(self, x: np.ndarray, j) -> float:
[…]

### 2023-01-04 src/gradient_algorithms.py

@@ -1,0 +1,171 @@
+from abc import ABC, abstractmethod
+from typing import List, Tuple
+
+import numpy as np
+
+from src.problems import Problem
+
+
+class GradientAlgorithm(ABC):
+    def __init__(self, problem: Problem):
+        self.problem = problem
+
+    @abstractmethod
+    def run(
+        self, x: np.ndarray, lambda_parameter: float, number_of_steps: int
+    ) -> Tuple[np.ndarray, List[float]]:
+        pass
+
+
+class ProximalStochasticGradientAlgorithm(GradientAlgorithm):
+    def __init__(self, problem: Problem, is_ergodic_mean: bool):
+        """
+        The Proximal Stochastic Gradient Algorithm.
+
+        :param problem: problem object
+        :param is_ergodic_mean: whether to use the ergodic mean for loss calculation
+        """
+        self.is_ergodic_mean = is_ergodic_mean
+        super().__init__(problem)
+
+    def generate_random_data_index(self) -> int:
+        """
+        Random choice of a datapoint index from (1, ..., n)
+        :return: randomly chosen index
+        """
+        return np.random.choice(self.problem.n)
+
+    def calculate_gamma_parameter(self, k: int):
+        """
+        gamma_k = n/(||A||^2 * sqrt (k+1))
+
+        :param k: step number
+        :return: gamma_k for the psga algorithm
+        """
+
+        return self.problem.n / (
+            np.linalg.norm(self.problem.a_matrix) ** 2 * np.sqrt(k + 1)
+        )
+
+    def run(
+        self, x: np.ndarray, lambda_parameter: float, number_of_steps: int
+    ) -> Tuple[np.ndarray, List[float]]:
+        """
+        The PSGA algorithm.
+
+        :param x: initial solution (number of dimension, 1)
+        :param lambda_parameter: scalar on g(x) of the minimization problem
+        :param number_of_steps: number of steps to run the algorithm
+        :return: optimised solution from PSGA and list of loss values at each iteration
+        """
+        gamma_normaliser = 0
+        x_bar_unnormalised = np.zeros(x.shape)
+        loss = [self.problem.loss(x, lambda_parameter)]
+        for k in range(number_of_steps):
+            i = self.generate_random_data_index()
+            gamma_parameter_k = self.calculate_gamma_parameter(k)
+            x = self.problem.proximity_operator(
+                x=(
+                    x  # (1, number of dimension)
+                    - gamma_parameter_k
[…]

### 2023-01-04 src/problems.py

@@ -1,0 +1,279 @@
+from __future__ import annotations
+
+import random
+from abc import ABC, abstractmethod
+
+import numpy as np
+from sklearn.datasets import make_moons
+
+
+class Problem(ABC):
+    def __init__(self, a_matrix: np.ndarray, y: np.ndarray):
+        """
+        A minimization problem.
+
+        :param a_matrix: the design matrix A (number of points, number of dimensions)
+        :param y: response vector (number of points, 1)
+        """
+        self.a_matrix = a_matrix
+        self.y = y
+
+    @property
+    def n(self):
+        return self.a_matrix.shape[0]
+
+    @property
+    def d(self):
+        return self.a_matrix.shape[1]
+
+    def a_matrix_i(self, i: int):
+        """
+        ith row of the A matrix
+            i = 1, ..., n
+
+        :param i: row index
+        :return: row vector of shape (1, number of dimensions)
+        """
+        return self.a_matrix[i : i + 1, :]
+
+    def a_matrix_j(self, j: int):
+        """
+        jth column of the A matrix
+            j = 1, ..., d
+
+        :param j: column index
+        :return: column vector of shape (number of points, 1)
+        """
+        return self.a_matrix[:, j : j + 1]
+
+    @staticmethod
+    @abstractmethod
+    def proximity_operator(
+        x: np.ndarray, gamma_parameter: float, lambda_parameter: float
+    ) -> np.ndarray:
+        """
+        Proximity operator attached to g(x) for the minimisation problem.
+
+        :param x: current solution (number of dimension, 1)
+        :param gamma_parameter: scalar dictating the step size
+        :param lambda_parameter: scalar on g(x) of the minimization problem
+        :return: prox(x) vector (number of dimensions, 1)
+        """
+        pass
+
+    @abstractmethod
+    def loss(self, x: np.ndarray, lambda_parameter: float) -> float:
+        """
+        Calculation of loss function with respect to x and lambda
+
+        :param x: current solution (number of dimension, 1)
+        :param lambda_parameter: scalar on g(x) of the minimization problem
+        :return: loss function evaluation at x with lambda
+        """
+        pass
+
+    @staticmethod
+    @abstractmethod
+    def generate(*args, **kwargs) -> Problem:
+        """
+        Instantiate a problem.
+        """
+        pass
+
+    @abstractmethod
[…]

### 2023-01-05 submission.tex
add figures for part 3
update algos and report

@@
 \maketitle
-\section*{Part 1}
+\section*{Part 1: Questions with multiple answers}
 \subsection*{1.1}
@@
 
-\section*{Part 2}
+\section*{Part 2: Theory on convex analysis and optimization}
 \subsection*{2.1}
@@
 
+We are considering the optimisation problem:
+
+\[\min_{\|\textbf{A}\textbf{x} - \textbf{b}\|_{\infty} \leq \epsilon} \frac{1}{2}\|\textbf{x}\|^2\]
+
+where $\epsilon > 0$, $\textbf{A} \in \mathbb{R}^{n\times d}$, $\textbf{b} \in \mathbb{R}^{n \times 1}$, and $\textbf{x} \in \mathbb{R}^{d \times 1}$.
+
+\subsubsection*{2.6.1}
+
+To compute the dual problem, we begin by reformulating the problem as:
+
+\[\min_{\textbf{x}}  f(\textbf{x}) + g(\textbf{Ax})\]
+
+where $f(x) = \frac{1}{2}\|\textbf{x}\|^2$ and $g(\textbf{Ax}) =  i_{[0, 1]}\left(\frac{1}{\epsilon}\|\textbf{A}\textbf{x} - \textbf{b}\|_{\infty}\right)$.
+
+Thus we have our primal problem:
+
+\[\min_{\textbf{x}}  \frac{1}{2}\|\textbf{x}\|^2 + i_{[0, 1]}\left(\frac{1}{\epsilon}\|\textbf{A}\textbf{x} - \textbf{b}\|_{\infty}\right)\]
+
+We know from the Fenchel-Rockafellar duality theory that the dual problem is formulated as:
+
+\[\min_{\textbf{u}}  g^*(\textbf{u}) + f^*(-\textbf{A}^*\textbf{u})\]
+
+where $\textbf{u} \in \mathbb{R}^{n \times 1}$, $\textbf{A}^* \in \mathbb{R}^{d\times n}$ the transpose of $\textbf{A}$, and $g^*$ and $f^*$ are the Fenchel conjugates of $g$ and $f$ respectively.
+
+The Fenchel conjugates are:
+
+\[g^*(\textbf{Ax}) = i^*_{[0, 1]}\left(\frac{1}{\epsilon}\|\textbf{A}\textbf{x} - \textbf{b}\|_{\infty}\right) = \frac{1}{\epsilon} \| \textbf{A}\textbf{x} - \textbf{b}\|_1\]
+
+\[f^*(\textbf{x}) =  \frac{1}{2}\|\textbf{x}\|^2\]
+
+We can substitute:
+
+\[\min_{\textbf{u}}  \frac{1}{\epsilon} \| \textbf{u} - \textbf{b}\|_1 +  \frac{1}{2}\|-\textbf{A}^*\textbf{u}\|^2\]
+
+Simplifying, we have our dual problem:
+
+\[\min_{\textbf{u}} \frac{1}{2}\textbf{u}^*\textbf{A}\textbf{A}^*\textbf{u} + \frac{1}{\epsilon} \| \textbf{u} - \textbf{b}\|_1\]
+
+\subsubsection*{2.6.2}
+
+To determine if strong duality holds, we use the qualification condition:
+
+\[\textbf{0}_n \in int \big(dom(g) - \textbf{A}dom(f)\big)\]
+
+that $\textbf{0}_n = \textbf{0} \in \mathbb{R}^{n \times 1}$ is in the interior of the intersection of  $dom(g)$, the domain of $g$, with $\textbf{A}dom(f)$, $\textbf{A}$ applied to the domain of $f$.
+
+We know that for $f(x) = \frac{1}{2}\|\textbf{x}\|^2$:
+
+\[\textbf{0}_d \in int \big(dom(f)\big)\]
+
+where $\textbf{0}_d = \textbf{0} \in \mathbb{R}^{d \times 1}$. Thus, knowing that $\textbf{A} \in \mathbb{R}^{n\times d}$ is a linear operator:
+
+\[\textbf{0}_n \in int \big(\textbf{A} dom(f)\big)\]
+
+Moreover, for $g(\textbf{u}) =  i_{[0, 1]}\left(\frac{1}{\epsilon}\|\textbf{u} - \textbf{b}\|_{\infty}\right)$:
+
+\[\textbf{0}_n \in dom(g)\]
+
+if $\|\textbf{b}\|_{\infty} \leq \epsilon$. Therefore:
+
+\[\textbf{0}_n \in int \big(dom(g)\big)\]
+
+if $\|\textbf{b}\|_{\infty} < \epsilon$.
+
+Under the condition that $\|\textbf{b}\|_{\infty} < \epsilon$, strong duality holds.
+This is because it will be the case that $\textbf{0}_n \in int \big(\textbf{A} dom(f)\big)$ and $\textbf{0}_n \in int \big(dom(g)\big)$ and so the same will hold for their intersection, $\textbf{0}_n \in int \big(dom(g) - \textbf{A}dom(f)\big)$ as required.
+
+\subsubsection*{2.6.3}
+
+The KKT conditions:
+
+\[\textbf{x} \in \partial f^*(-\textbf{A}^*\textbf{u}) \text{ and } \textbf{Ax} \in \partial g^*(\textbf{u})\]
+
+where $\partial f^*$ and $\partial g^*$ are the subgradients of $f^*$ and $g^*$ respectively.
+
+We know that:
+
+\[\partial f^*(-\textbf{A}^*\textbf{u}) = \partial \frac{1}{2}\|-\textbf{A}^*\textbf{u}\|^2 = -\textbf{A}^*\textbf{u}\]
+
+Moreover:
+
+\[\partial g^*(\textbf{u}) = \partial\frac{1}{\epsilon} \| \textbf{u} - \textbf{b}\|_1 \]
+
+and so
+
+\[\left(\partial\| \textbf{u} - \textbf{b}\|_1\right)_i = \begin{cases}
+      \frac{1}{\epsilon}, &  \text{if } | u_i - b_i| > 0 \\
+      -\frac{1}{\epsilon}, &  \text{if } | u_i - b_i| < 0 \\
+      [-\frac{1}{\epsilon}, \frac{1}{\epsilon}], &  \text{if } | u_i - b| = 0 \\
+   \end{cases} \\
+\]
+
+where $\left(\partial\| \textbf{u} - \textbf{b}\|_1\right)_i$ is the $i^{th}$ element of $\partial\| \textbf{u} - \textbf{b}\|_1 \in \mathbb{R}^{n \times 1}$
+
+Thus, our KKT conditions state that:
+
+\[\textbf{x} = -\textbf{A}^*\textbf{u}\]
+
+and
+
+\[(\textbf{Ax})_i \in \begin{cases}
+      \frac{1}{\epsilon}, &  \text{if } | u_i - b_i| > 0 \\
+      -\frac{1}{\epsilon}, &  \text{if } | u_i - b_i| < 0 \\
+      [-\frac{1}{\epsilon}, \frac{1}{\epsilon}], &  \text{if } | u_i - b| = 0 \\
+   \end{cases} \\\]
+
+where $(\textbf{Ax})_i$ is the $i^{th}$ element of $\textbf{Ax}\in \mathbb{R}^{n \times 1}$
+
+\subsubsection*{2.6.4}
+
+To derive a rate of convergence on the primal iterates from the applications of FISTA (Fast Iterative Shrinkage Threshold Algorithm) on the  dual problem when strong duality holds, the distance to the primal solution is bounded by the dual objective values:
+
+\[\frac{2}{\mu} \|\textbf{x}_k - \bar{\textbf{x}}\|^2 \leq \Psi(\textbf{u}_k) - \Psi(\bar{\textbf{u}})\]
+
+where $\bar{\textbf{x}} \in \mathbb{R}^{d \times 1}$ and $\bar{\textbf{u}} \in \mathbb{R}^{n \times 1}$ are minimizers for the primal and dual solution respectively, $\textbf{x}_k \in \mathbb{R}^{d \times 1}$ and $\textbf{u}_k \in \mathbb{R}^{n \times 1}$ are respectively the primal and dual solutions for the $k^{th}$ step of FISTA, and $\mu$ is the coefficient of strong convexity for $f$.
+For our problem, $\mu = 1$.
+Moreover, $\Psi(\textbf{u})=\frac{1}{2}\textbf{u}^*\textbf{A}\textbf{A}^*\textbf{u} + \frac{1}{\epsilon} \| \textbf{u} - \textbf{b}\|_1$, our dual objective.
+
+We also know for FISTA that (proof in lecture):
+
+\[\Psi(\textbf{u}_k) - \Psi(\bar{\textbf{u}}) \leq \frac{\|\textbf{u}_0 - \bar{\textbf{u}}\|^2}{2 \gamma \left(t_{k-1}}\right)^2\]
+
+where $\textbf{u}_0 \in \mathbb{R}^{n \times 1}$ is the initial starting point for FISTA, and $0 < \gamma \leq \frac{1}{L}$  and $t_{k-1}$ are both scalars dictating step sizes in FISTA ($t_{k-1}$ is for step $k-1}$).
+$L$ is the Lipschitz smoothness of $f$.
+For our problem, $L=1$.
+
+Thus:
+
+\[ \|\textbf{x}_k - \bar{\textbf{x}}\|^2 \leq \frac{\|\textbf{u}_0 - \bar{\textbf{u}}\|^2}{4 \gamma \left(t_{k-1}}\right)^2\]
+
+Simplifying:
+
+\[ \|\textbf{x}_k - \bar{\textbf{x}}\| \leq \frac{\|\textbf{u}_0 - \bar{\textbf{u}}\|}{2  \sqrt{\gamma}t_{k-1}}\]
+
+Two choices for $t_k$ in FISTA are:
+
+\begin{itemize}
+\setlength{\itemindent}{2em}
+\item $t_k = \frac{1+\sqrt{1+ 4 (t_{k-1})^2}}{2}$ where $\frac{1}{t_{k-1}} \leq \frac{2}{k+1}$
+\item $t_k = \frac{k+a}{a}$ with $a \geq 2$ where $\frac{1}{t_{k-1}} \leq \frac{a}{k+1}$
+\end{itemize}
+
+where $t_0  =1 $
+
+In both cases the rate of convergence of the primal iterates when applying FISTA on the dual problem:
+
+\[ \|\textbf{x}_k - \bar{\textbf{x}}\| \leq \frac{a\|\textbf{u}_0 - \bar{\textbf{u}}\|}{\sqrt{\gamma}(k+1)}\]
+
+where $a = 2$ if $t_k = \frac{1+\sqrt{1+ 4 (t_{k-1})^2}}{2}$. In other words, the rate is $\mathcal{O}(\frac{1}{k})$.
+
 \newpage
-\section*{Part 3}
+\section*{Part 3: Solving the Lasso problem}
 
@@
 \newpage
-\section*{Part 4}
+\subsection*{3.3}
+
+Plotting the objective function values vs the number of iterations for both the algorithms:
+
+\begin{figure}[h]
+\centering
+\begin{minipage}{.5\textwidth}
+  \centering
+\includegraphics[scale=0.4]{outputs/part_3/psga-loss}
+\caption{PSGA}
+\label{fig:}
+\end{minipage}%
+\begin{minipage}{.5\textwidth}
+  \centering
+\includegraphics[scale=0.4]{outputs/part_3/rcpga-loss}
+\caption{RCPGA}
+\label{fig:}
+\end{minipage}
+\end{figure}
+
+The corresponding solution compared to the actual sparse vector for both the algorithms:
+
+\begin{figure}[h]
+\centering
+\begin{minipage}{.5\textwidth}
+  \centering
+\includegraphics[scale=0.4]{outputs/part_3/psga-x}
+\caption{PSGA}
+\label{fig:}
+\end{minipage}%
+\begin{minipage}{.5\textwidth}
+  \centering
+\includegraphics[scale=0.4]{outputs/part_3/rcpga-x}
+\caption{RCPGA}
+\label{fig:}
+\end{minipage}
+\end{figure}
+
+\newpage
+\section*{Part 4: Support Vector Machines}
 
@@
 
+\subsection*{4.2}
+
+\newpage
+\subsection*{4.3}
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.25]{outputs/part_4/initial-contour}
+\caption{$\alpha^{(0)}$ Contour Plot}
+\label{fig:}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\begin{minipage}{.5\textwidth}
+  \centering
+\includegraphics[scale=0.25]{outputs/part_4/rcpga-contour}
+\caption{RCPGA Contour Plot}
+\label{fig:}
+\end{minipage}%
+\begin{minipage}{.5\textwidth}
+  \centering
+\includegraphics[scale=0.25]{outputs/part_4/fista-contour}
+\caption{FISTA Contour Plot}
+\label{fig:}
+\end{minipage}
+\end{figure}
+
+\subsection*{4.4}
+
+\begin{figure}[h]
+\centering
+\begin{minipage}{.5\textwidth}
+  \centering
+\includegraphics[scale=0.4]{outputs/part_4/rcpga-loss}
+\caption{RCPGA Loss}
+\label{fig:}
+\end{minipage}%
+\begin{minipage}{.5\textwidth}
+  \centering
+\includegraphics[scale=0.4]{outputs/part_4/fista-loss}
+\caption{FISTA Loss}
+\label{fig:}
+\end{minipage}
+\end{figure}
+
 \end{document}

### 2023-01-05 helpers.py

@@ -1,0 +1,20 @@
+from typing import List
+
+import matplotlib.pyplot as plt
+
+
+def plot_loss(
+    losses: List[List[float]], labels: List[str], algorithm: str, save_path: str
+):
+    fig = plt.figure()
+    fig.set_figwidth(10)
+    fig.set_figheight(5)
+    for i, loss in enumerate(losses):
+        plt.plot(loss, label=labels[i])
+    if len(labels) > 1:
+        plt.legend()
+    plt.xlabel("Iteration")
+    plt.ylabel("Loss")
+    plt.title(f"Loss ({algorithm})")
+    plt.savefig(save_path, bbox_inches="tight")
+    plt.close()

### 2023-01-05 main.py

@@ -1,3 +1,0 @@
-from part_3 import part_3
-
-part_3()

### 2023-01-05 part_3.py

@@ -1,32 +1,14 @@
 import os
-from typing import List
 
 import matplotlib.pyplot as plt
 import numpy as np
 
 from constants import DEFAULT_SEED, OUTPUTS_FOLDER
+from helpers import plot_loss
 from src.gradient_algorithms import (
     ProximalStochasticGradientAlgorithm,
-    RandomizedCoordinateProximalGradientAlgorithm,
-)
+    RandomizedCoordinateProximalGradientAlgorithm)
 from src.problems import SparseProblem
-
-
-def plot_loss(
-    losses: List[List[float]], labels: List[str], algorithm: str, save_path: str
-):
-    fig = plt.figure()
-    fig.set_figwidth(10)
-    fig.set_figheight(5)
-    for i, loss in enumerate(losses):
-        plt.plot(loss, label=labels[i])
-    if len(labels) > 1:
-        plt.legend()
-    plt.xlabel("Iteration")
-    plt.ylabel("Loss")
-    plt.title(f"Loss ({algorithm})")
-    plt.savefig(save_path, bbox_inches="tight")
-    plt.close()
 
 
 def plot_dimensions(
@@ -86,8 +68,8 @@ def part_3():
         std=0.06,
     )
     x0 = np.random.randn(sparse_problem.d).reshape(-1, 1)
-    lambda_parameter = 10
-    number_of_steps = 1000
+    lambda_parameter = 5e-2
+    number_of_steps = int(1e6)
 
     # Proximal Stochastic Gradient Algorithm
     proximal_stochastic_gradient_algorithm = ProximalStochasticGradientAlgorithm(
@@ -127,6 +109,7 @@ def part_3():
     )
 
     # Randomized Coordinate Proximal Gradient Algorithm
+    number_of_steps = int(1e4)
     randomized_coordinate_proximal_gradient_algorithm = (
         RandomizedCoordinateProximalGradientAlgorithm(sparse_problem)
     )

### 2023-01-05 part_4.py

@@ -1,0 +1,96 @@
+import os
+from typing import List
+
+import matplotlib.pyplot as plt
+import numpy as np
+
+from constants import DEFAULT_SEED, OUTPUTS_FOLDER
+from helpers import plot_loss
+from src.gradient_algorithms import (
+    FastIterativeShrinkageThresholdAlgorithm,
+    RandomizedCoordinateProjectedGradientAlgorithm)
+from src.problems import HalfMoonsProblem
+
+
+def plot_contour(half_moons_problem, x, algorithm, save_path):
+    x_ticks = np.linspace(
+        np.min(half_moons_problem.x[:, 0]) - 1,
+        np.max(half_moons_problem.x[:, 0]) + 1,
+        100,
+    )
+    y_ticks = np.linspace(
+        np.min(half_moons_problem.x[:, 1]) - 1,
+        np.max(half_moons_problem.x[:, 1]) + 1,
+        100,
+    )
+
+    x1_grid, x2_grid = np.meshgrid(x_ticks, y_ticks)
+    x_grid = np.stack((x1_grid.flatten(), x2_grid.flatten())).T
+
+    y_grid = half_moons_problem.predict(x.T, x_grid).reshape(-1)
+    y = half_moons_problem.y.reshape(-1)
+    fig = plt.figure()
+    fig.set_figwidth(12)
+    fig.set_figheight(12)
+    plt.contourf(x1_grid, x2_grid, y_grid.reshape(x1_grid.shape))
+    for label in set(y):
+        idx = y == label
+        plt.scatter(
+            half_moons_problem.x[idx, 0],
+            half_moons_problem.x[idx, 1],
+            label=label,
+        )
+    plt.legend()
+    plt.title(f"Contour Plot ({algorithm})")
+    plt.savefig(save_path, bbox_inches="tight")
+    plt.close()
+
+
+def part_4():
+    np.random.seed(DEFAULT_SEED)
+
+    # Part 3
+
+    part_4_output_folder = os.path.join(OUTPUTS_FOLDER, "part_4")
+    if not os.path.exists(part_4_output_folder):
+        os.makedirs(part_4_output_folder)
+
+    half_moons_problem = HalfMoonsProblem.generate(
+        number_of_samples=200,
+        noise=0.05,
+        sigma=1,
+        random_state=0,
+    )
+    lambda_parameter = 1e-1
+    x0 = np.random.uniform(
+        low=0,
+        high=1 / (lambda_parameter * half_moons_problem.n),
[…]

### 2023-01-05 src/gradient_algorithms.py

@@ -3,7 +3,7 @@ from typing import List, Tuple
 
 import numpy as np
 
-from src.problems import Problem
+from src.problems import HalfMoonsProblem, Problem
 
 
 class GradientAlgorithm(ABC):
@@ -138,28 +138,63 @@ class RandomizedCoordinateProximalGradientAlgorithm(GradientAlgorithm):
         return x, loss
 
 
-class FastIterativeShrinkageThresholdAlgorithm(GradientAlgorithm):
-    def __init__(self, problem: Problem):
+class RandomizedCoordinateProjectedGradientAlgorithm(
+    RandomizedCoordinateProximalGradientAlgorithm
+):
+    def __init__(self, problem: HalfMoonsProblem):
         super().__init__(problem)
+        self.problem = problem
+
+    def run(
+        self, x, lambda_parameter, number_of_steps
+    ) -> Tuple[np.ndarray, List[float]]:
+        """
+        The RCPGA (projection) algorithm.
+
+        :param x: initial solution (number of dimension, 1)
+        :param lambda_parameter: scalar on g(x) of the minimization problem
+        :param number_of_steps: number of steps to run the algorithm
+        :return: optimised solution from RCPGA and list of loss values at each iteration
+        """
+        loss = [self.problem.loss(x, lambda_parameter)]
+        for k in range(number_of_steps):
+            j = self.generate_random_dimension_index()
+            gamma_parameter_j = self.calculate_gamma_parameter(j)
+            x[j] = self.problem.projection_operator(
+                x=(
+                    x[j]
+                    - (gamma_parameter_j / self.problem.n) * self.problem.grad_f_j(x, j)
+                ),
+                lambda_parameter=lambda_parameter,
+            )
+            loss.append(self.problem.loss(x, lambda_parameter))
+        return x, loss
+
+
+class FastIterativeShrinkageThresholdAlgorithm(GradientAlgorithm):
+    def __init__(self, problem: HalfMoonsProblem):
+        super().__init__(problem)
+        self.problem = problem
 
     @staticmethod
     def calculate_new_t(t):
         return (1 + np.sqrt(1 + 4 * t**2)) / 2
 
-    def calculate_gamma(self) -> float:
-        return 2 / (np.linalg.norm(self.problem.a_matrix) ** 2)
+    def calculate_gamma_parameter(self) -> float:
+        return 2 * self.problem.n / (np.linalg.norm(self.problem.a_matrix) ** 2)
 
     def run(
[…]

### 2023-01-05 src/problems.py

@@ -129,7 +129,9 @@ class SparseProblem(Problem):
         super().__init__(a_matrix, y)
 
     @staticmethod
-    def proximity_operator(x, gamma_parameter, lambda_parameter):
+    def proximity_operator(
+        x: np.ndarray, gamma_parameter: float, lambda_parameter: float
+    ) -> np.ndarray:
         """
         The soft thresholding operator, the proximity operator for the absolute value function:
             lambda*||·||
@@ -142,7 +144,7 @@ class SparseProblem(Problem):
         rho = gamma_parameter * lambda_parameter
         return np.sign(x - rho) * np.maximum(0, np.abs(x) - rho)
 
-    def loss(self, x: np.ndarray, lambda_parameter: float):
+    def loss(self, x: np.ndarray, lambda_parameter: float) -> float:
         """
 
         :param x: current solution (number of dimension, 1)
@@ -209,31 +211,68 @@ class SparseProblem(Problem):
 
 class HalfMoonsProblem(Problem):
     def __init__(self, x: np.ndarray, y: np.ndarray, sigma: float):
+        """
+        Classification dataset
+
+        :param x: design_matrix of training points (number of training points, 2)
+        :param y: the response label -1 or 1 (number of training points, 1)
+        :param sigma: noise parameter during data generation process
+        """
         self.x = x
         self.sigma = sigma
-        self.gram_matrix = self.calculate_gram(x)
+        self.gram_matrix = self.calculate_gram(sigma, x)
         super().__init__(
             a_matrix=np.multiply(y @ y.T, self.gram_matrix),
             y=y,
         )
 
     @staticmethod
-    def kernel(x, y, sigma):
-        diff = x - y
-        return np.exp(-diff.T @ diff / (2 * sigma**2))
+    def kernel(x: np.ndarray, y: np.ndarray, sigma: float) -> float:
+        """
+        Square exponential kernel.
 
-    def calculate_gram(self, x):
-        n = self.x.shape[0]
-        m = x.shape[0]
+        :param x: data vector (2, 1)
+        :param y: data vector (2, 1)
+        :param sigma: kernel length scale
+        :return: kernel evaluation k(x, y)
+        """
+        diff = x - y
+        return np.exp(-diff.T @ diff / (2 * sigma**2)).item()
+
+    @staticmethod
+    def calculate_gram(sigma: float, x: np.ndarray, y: np.ndarray = None) -> np.ndarray:
+        """
[…]

### 2023-01-06 constants.py
implement fista

@@ -1,3 +1,4 @@
 OUTPUTS_FOLDER = "outputs"
 
 DEFAULT_SEED = 0
+FEATURE_THRESHOLD = 1e-3

### 2023-01-06 part_3.py

@@ -3,7 +3,7 @@ import os
 import matplotlib.pyplot as plt
 import numpy as np
 
-from constants import DEFAULT_SEED, OUTPUTS_FOLDER
+from constants import DEFAULT_SEED, FEATURE_THRESHOLD, OUTPUTS_FOLDER
 from helpers import plot_loss
 from src.gradient_algorithms import (
     ProximalStochasticGradientAlgorithm,
@@ -36,8 +36,8 @@ def plot_dimensions(
         label="$x^*$",
     )
     ax.stem(
-        np.arange(d)[np.abs(x) > 0],
-        x[np.abs(x) > 0],
+        np.arange(d)[np.abs(x) > FEATURE_THRESHOLD],
+        x[np.abs(x) > FEATURE_THRESHOLD],
         label=r"$x_{\tau, \lambda}$",
         linefmt="k:",
         markerfmt="k^",
@@ -46,7 +46,7 @@ def plot_dimensions(
     ax.set_xlim([-10 + 0, d + 10])
     ax.set_ylim([-1.1, 1.1])
     ax.set_xlabel("index $i$")
-    ax.set_ylabel("$x_i$")
+    ax.set_ylabel(f"$x_i$ ({FEATURE_THRESHOLD=})")
     ax.legend()
     plt.title(f"Sparse Solution vs Actual ({algorithm})")
     plt.savefig(save_path, bbox_inches="tight")
@@ -68,8 +68,14 @@ def part_3():
         std=0.06,
     )
     x0 = np.random.randn(sparse_problem.d).reshape(-1, 1)
-    lambda_parameter = 5e-2
-    number_of_steps = int(1e6)
+    plot_dimensions(
+        x=x0,
+        x_sparse=sparse_problem.x_sparse,
+        algorithm=f"Initialised Vector (x0)",
+        save_path=os.path.join(part_3_output_folder, f"initial-x"),
+    )
+    lambda_parameter = 3e-1
+    number_of_steps = int(5e5)
 
     # Proximal Stochastic Gradient Algorithm
     proximal_stochastic_gradient_algorithm = ProximalStochasticGradientAlgorithm(
@@ -85,7 +91,7 @@ def part_3():
         x=x_psga,
         x_sparse=sparse_problem.x_sparse,
         algorithm=f"Proximal Stochastic Gradient Algorithm (x), {lambda_parameter=}",
-        save_path=os.path.join(part_3_output_folder, "psga-x"),
[…]

### 2023-01-06 part_4.py

@@ -13,14 +13,18 @@ from src.problems import HalfMoonsProblem
 
 
 def plot_contour(half_moons_problem, x, algorithm, save_path):
+    x_min = np.min(half_moons_problem.x[:, 0]) - 1
+    x_max = np.max(half_moons_problem.x[:, 0]) + 1
+    y_min = np.min(half_moons_problem.x[:, 1]) - 1
+    y_max = np.max(half_moons_problem.x[:, 1]) + 1
     x_ticks = np.linspace(
-        np.min(half_moons_problem.x[:, 0]) - 1,
-        np.max(half_moons_problem.x[:, 0]) + 1,
+        x_min,
+        x_max,
         100,
     )
     y_ticks = np.linspace(
-        np.min(half_moons_problem.x[:, 1]) - 1,
-        np.max(half_moons_problem.x[:, 1]) + 1,
+        y_min,
+        y_max,
         100,
     )
 
@@ -30,8 +34,8 @@ def plot_contour(half_moons_problem, x, algorithm, save_path):
     y_grid = half_moons_problem.predict(x.T, x_grid).reshape(-1)
     y = half_moons_problem.y.reshape(-1)
     fig = plt.figure()
-    fig.set_figwidth(12)
-    fig.set_figheight(12)
+    fig.set_figwidth(2 * (x_max - x_min))
+    fig.set_figwidth(2 * (y_max - y_min))
     plt.contourf(x1_grid, x2_grid, y_grid.reshape(x1_grid.shape))
     for label in set(y):
         idx = y == label
@@ -49,46 +53,60 @@ def plot_contour(half_moons_problem, x, algorithm, save_path):
 def part_4():
     np.random.seed(DEFAULT_SEED)
 
-    # Part 3
-
+    # Part 4
     part_4_output_folder = os.path.join(OUTPUTS_FOLDER, "part_4")
     if not os.path.exists(part_4_output_folder):
         os.makedirs(part_4_output_folder)
 
+    number_of_samples = 200
+    x0 = np.random.randn(number_of_samples).reshape(-1, 1)
+    sigma = 0.4
+    lambda_parameter = 10
+    number_of_steps = int(2e5)
     half_moons_problem = HalfMoonsProblem.generate(
-        number_of_samples=200,
+        number_of_samples=number_of_samples,
         noise=0.05,
-        sigma=1,
+        sigma=sigma,
         random_state=0,
     )
-    lambda_parameter = 1e-1
-    x0 = np.random.uniform(
-        low=0,
[…]

### 2023-01-06 src/gradient_algorithms.py

@@ -44,7 +44,7 @@ class ProximalStochasticGradientAlgorithm(GradientAlgorithm):
         """
 
         return self.problem.n / (
-            np.linalg.norm(self.problem.a_matrix) ** 2 * np.sqrt(k + 1)
+            np.linalg.norm(self.problem.a_matrix, ord=np.inf) ** 2 * np.sqrt(k + 1)
         )
 
     def run(
@@ -107,9 +107,7 @@ class RandomizedCoordinateProximalGradientAlgorithm(GradientAlgorithm):
         :param j: dimension index
         :return: gamma_j for the rcpga algorithm
         """
-        return self.problem.n / (
-            self.problem.a_matrix_j(j).T @ self.problem.a_matrix_j(j)
-        )
+        return self.problem.n / (np.linalg.norm(self.problem.a_matrix_j(j)) ** 2)
 
     def run(
         self, x, lambda_parameter, number_of_steps
@@ -141,34 +139,14 @@ class RandomizedCoordinateProximalGradientAlgorithm(GradientAlgorithm):
 class RandomizedCoordinateProjectedGradientAlgorithm(
     RandomizedCoordinateProximalGradientAlgorithm
 ):
-    def __init__(self, problem: HalfMoonsProblem):
-        super().__init__(problem)
-        self.problem = problem
-
-    def run(
-        self, x, lambda_parameter, number_of_steps
-    ) -> Tuple[np.ndarray, List[float]]:
+    def calculate_gamma_parameter(self, j: int) -> float:
         """
-        The RCPGA (projection) algorithm.
+        gamma = 1/(||a_j||^2)
 
-        :param x: initial solution (number of dimension, 1)
-        :param lambda_parameter: scalar on g(x) of the minimization problem
-        :param number_of_steps: number of steps to run the algorithm
-        :return: optimised solution from RCPGA and list of loss values at each iteration
+        :param j: dimension index
+        :return: gamma_j for the rcpga algorithm
         """
-        loss = [self.problem.loss(x, lambda_parameter)]
-        for k in range(number_of_steps):
-            j = self.generate_random_dimension_index()
-            gamma_parameter_j = self.calculate_gamma_parameter(j)
-            x[j] = self.problem.projection_operator(
-                x=(
-                    x[j]
-                    - (gamma_parameter_j / self.problem.n) * self.problem.grad_f_j(x, j)
-                ),
-                lambda_parameter=lambda_parameter,
-            )
[…]

### 2023-01-06 src/problems.py

@@ -20,10 +20,12 @@ class Problem(ABC):
 
     @property
     def n(self):
+        #  number of points
         return self.a_matrix.shape[0]
 
     @property
     def d(self):
+        #  number of dimensions
         return self.a_matrix.shape[1]
 
     def a_matrix_i(self, i: int):
@@ -182,10 +184,10 @@ class SparseProblem(Problem):
         """
         Generate xs vectors with entries in [0.5, 1] and [-1, -0.5] respectively.
 
-        :param number_of_samples:
-        :param number_of_dimensions:
-        :param sparsity:
-        :param std:
+        :param number_of_samples: number of samples to generate
+        :param number_of_dimensions: number of dimensions to generate
+        :param sparsity: the number of non-zero elements in x_sparse
+        :param std: standard deviation parameterizing epsilon, the noise added to the observed response y
         :return: spare problem
         """
 
@@ -216,7 +218,7 @@ class HalfMoonsProblem(Problem):
 
         :param x: design_matrix of training points (number of training points, 2)
         :param y: the response label -1 or 1 (number of training points, 1)
-        :param sigma: noise parameter during data generation process
+        :param sigma: kernel length scale
         """
         self.x = x
         self.sigma = sigma
@@ -261,55 +263,78 @@ class HalfMoonsProblem(Problem):
                 )
         return gram_matrix
 
-    @staticmethod
     def proximity_operator(
-        x: np.ndarray, gamma_parameter: float, lambda_parameter: float
+        self, x: np.ndarray, gamma_parameter: float, lambda_parameter: float
     ) -> np.ndarray:
         """
+        Projection to the convex set [0, 1/(lambda*n)]^n
 
         :param x: current solution (number of training points, 1)
-        :param gamma_parameter: scalar dictating the step size
-        :param lambda_parameter: scalar on g(x) of the minimization problem
-        :return: soft_threshold(x) vector (number of dimensions, 1)
-        """
-        rho = gamma_parameter / lambda_parameter
-        out = np.copy(x)
-        mask_1 = x > 1
-        out[mask_1] = 0
-
-        mask_2 = x < (1 - rho)
-        out[mask_2] = rho
-
-        out[~(mask_1 | mask_2)] = 1 - x[~(mask_1 | mask_2)]
-        return out
-
-    def projection_operator(self, x: np.ndarray, lambda_parameter: float) -> np.ndarray:
-        """
-
[…]

### 2023-01-06 part_3.py
add type hints

@@ -7,7 +7,8 @@ from constants import DEFAULT_SEED, FEATURE_THRESHOLD, OUTPUTS_FOLDER
 from helpers import plot_loss
 from src.gradient_algorithms import (
     ProximalStochasticGradientAlgorithm,
-    RandomizedCoordinateProximalGradientAlgorithm)
+    RandomizedCoordinateProximalGradientAlgorithm,
+)
 from src.problems import SparseProblem
 
 

### 2023-01-06 part_4.py

@@ -8,11 +8,14 @@ from constants import DEFAULT_SEED, OUTPUTS_FOLDER
 from helpers import plot_loss
 from src.gradient_algorithms import (
     FastIterativeShrinkageThresholdAlgorithm,
-    RandomizedCoordinateProjectedGradientAlgorithm)
+    RandomizedCoordinateProjectedGradientAlgorithm,
+)
 from src.problems import HalfMoonsProblem
 
 
-def plot_contour(half_moons_problem, x, algorithm, save_path):
+def plot_contour(
+    half_moons_problem: HalfMoonsProblem, x: np.ndarray, algorithm: str, save_path: str
+) -> None:
     x_min = np.min(half_moons_problem.x[:, 0]) - 1
     x_max = np.max(half_moons_problem.x[:, 0]) + 1
     y_min = np.min(half_moons_problem.x[:, 1]) - 1

### 2023-01-06 src/gradient_algorithms.py

@@ -3,7 +3,7 @@ from typing import List, Tuple
 
 import numpy as np
 
-from src.problems import HalfMoonsProblem, Problem
+from src.problems import HalfMoonsProblem, Problem, SparseProblem
 
 
 class GradientAlgorithm(ABC):
@@ -18,7 +18,7 @@ class GradientAlgorithm(ABC):
 
 
 class ProximalStochasticGradientAlgorithm(GradientAlgorithm):
-    def __init__(self, problem: Problem, is_ergodic_mean: bool):
+    def __init__(self, problem: SparseProblem, is_ergodic_mean: bool):
         """
         The Proximal Stochastic Gradient Algorithm.
 
@@ -139,6 +139,9 @@ class RandomizedCoordinateProximalGradientAlgorithm(GradientAlgorithm):
 class RandomizedCoordinateProjectedGradientAlgorithm(
     RandomizedCoordinateProximalGradientAlgorithm
 ):
+    def __init__(self, problem: HalfMoonsProblem):
+        super().__init__(problem)
+
     def calculate_gamma_parameter(self, j: int) -> float:
         """
         gamma = 1/(||a_j||^2)

### 2023-01-06 submission.tex
implement fista
update report

@@
 
+The initial vector passed to the algorithms:
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.35]{outputs/part_3/initial-x}
+\caption{$x_0$ vs Sparse Vector}
+\label{fig:}
+\end{figure}
+
 Plotting the objective function values vs the number of iterations for both the algorithms:
@@
   \centering
-\includegraphics[scale=0.4]{outputs/part_3/psga-loss}
-\caption{PSGA}
+\includegraphics[scale=0.3]{outputs/part_3/psga-loss}
+\caption{PSGA Loss}
 \label{fig:}
@@
   \centering
-\includegraphics[scale=0.4]{outputs/part_3/rcpga-loss}
-\caption{RCPGA}
+\includegraphics[scale=0.3]{outputs/part_3/rcpga-loss}
+\caption{RCPGA Loss}
 \label{fig:}
@@
   \centering
-\includegraphics[scale=0.4]{outputs/part_3/psga-x}
-\caption{PSGA}
+\includegraphics[scale=0.35]{outputs/part_3/psga-x}
+\caption{PSGA Solution vs Sparse Vector}
 \label{fig:}
@@
   \centering
-\includegraphics[scale=0.4]{outputs/part_3/rcpga-x}
-\caption{RCPGA}
+\includegraphics[scale=0.35]{outputs/part_3/rcpga-x}
+\caption{RCPGA Solution vs Sparse Vector}
 \label{fig:}
@@
 We are given the primal problem:
+\[\min_{\textbf{w} \in \mathcal{H}} \frac{1}{\lambda n} \sum_{i=1}^{n} (1-y_i \langle \textbf{w}, \Lambda(\textbf{x}_i)\rangle)_{+} + \frac{1}{2}\| \textbf{w}\|^2\]
 
-\[\min_{w \in \mathcal{H}} \frac{\lambda}{n} \sum_{i=1}^{n} (1-y_i \langle w, \Lambda(x_i)\rangle)_{+} + \frac{\lambda}{2}\| w\|^2\]
+where $\mathcal{H}$ is a Hilbert space.
 
-We can express in the form:
-
-\[\min_{w \in \mathcal{H}} g(w) + f(w)\]
+The primal problem can be expressed the form:
+\[\min_{w \in \mathcal{H}} f(\textbf{w}) + g(\textbf{A}\textbf{w})\]
 
 where:
+\[f(\textbf{w}) = \frac{1}{2}\| \textbf{w}\|^2\]
 
-\[g(w) = \frac{\lambda}{n} \sum_{i=1}^{n} (1-y_i \langle w, \Lambda(x_i)\rangle)_{+}\]
+and
+\[g(\textbf{A}\textbf{w}) = \frac{1}{\lambda n} \sum_{i=1}^{n} (1-y_i \langle \textbf{w}, \Lambda(\textbf{x}_i)\rangle)_{+}\]
+
+where $\textbf{A}_i = \Lambda(\textbf{x}_i) \in \mathcal{H}$ our Hilbert space.
+
+This corresponds to the dual problem of the form:
+\[\min_{\textbf{u} \in \mathbb{R}^{n}} f^*(-\textbf{A}^*\textbf{u}) + g^*(\textbf{u})\]
+
+where:
+\[f^*(-\textbf{A}^*\textbf{u}) = \frac{1}{2}\| -\textbf{A}^*\textbf{u}\|^2 = \frac{1}{2} \textbf{u}^*\textbf{A} \textbf{A}^*\textbf{u}\]
+
+and using the kernel defined by $\mathcal{H}$, we can substitute $\textbf{A} \textbf{A}^*$ with the gram matrix $\textbf{K}$:
+\[f^*(-\textbf{A}^*\textbf{u}) = \frac{1}{2} \textbf{u}^*\textbf{K}\textbf{u}\]
+
+Moreover,
+\[g^*(\textbf{u}) = - \langle \textbf{y}, \textbf{u}\rangle + \sum_{i=1}^{n} i_{\left[0, \frac{1}{\lambda n}\right]}(\textbf{u}_i)\]
+
+Defining $\alpha_i = u_i y_i$ and know that $y_i y_i = 1$ because $y_i \in \{-1, 1\}$ we can see that:
+\[f^*(-\textbf{A}^*\textbf{u}) = \frac{1}{2} \alpha^*\textbf{K}_y\alpha\]
+
+and
+\[g^*(\textbf{u}) = - \langle \textbf{1}_n, \alpha \rangle + \sum_{i=1}^{n} i_{\left[0, \frac{1}{\lambda n}\right]}(\alpha_i)\]
+
+where $(\textbf{K}_y)_{i, j} = y_i \textbf{K}_{i, j} y_j$.
+
+Combining, we have our dual problem:
+\[\min_{\alpha \in \mathbb{R}^n} \frac{1}{2} \alpha^*\textbf{K}_y\alpha - \langle \textbf{1}_n, \alpha \rangle + \sum_{i=1}^{n} i_{\left[0, \frac{1}{\lambda n}\right]}(\alpha_i)\]
+
+\newpage
+\subsection*{4.1}
+
+The Fast Iterative Shrinkage Threshold Algorithm (FISTA):
+
+\[\alpha_{k+1} = prox_{\gamma g^*(\cdot)}\bigg(\nu_k + \gamma \textbf{A}_y \nabla f^*(-\textbf{A}_y^* \alpha_{k})\bigg)\]
+
+\[\nu_{k+1} = \alpha_{k} + \frac{t_k - 1}{t_{k+1}}(\alpha_{k+1}-\alpha_{k})\]
+
+where $t_k = \frac{1+\sqrt{1+ 4 (t_{k-1})^2}}{2}$ or $t_k = \frac{k+a}{a}$ with $a \geq 2$. For our implementation we chose the first definition for $t_k$.
+
+For our dual problem:
+
+\[\nabla f^*(-\textbf{A}_y^* \alpha_{k})) = -\textbf{A}_y^* \alpha_{k}\]
 
@@
 
-\[f(w) = \frac{\lambda}{2}\| w\|^2\]
+\[prox_{\gamma g^*(\cdot)}(\omega) = \begin{cases}
+      0, &  \text{if } \omega < 0\\
+      \omega, &  \text{if } 0 \leq \omega \leq \frac{1}{\gamma n} \\
+      \frac{1}{\gamma n}, & \text{if } \frac{1}{\gamma n} < \omega \\
+   \end{cases}\]
 
-The corresponding dual problem:
+With FISTA implemented, we can plot the dual objective function:
 
-\[\min_{\alpha \in \mathbb{R}^n} \frac{1}{2} \langle \mathbf{K}_y \alpha, \alpha \rangle - \langle \textbf{1}_n, \alpha\rangle + \sum_{i=1}^{n} i_{\left[0, \frac{1}{\lambda n}\right]}(\alpha_i)\]
-
-We can express in the form:
-
-\[\min_{\alpha \in \mathbb{R}^n} g^*(\alpha) + f^*(\alpha)\]
-
-\subsection*{4.1}
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.3]{outputs/part_4/fista-loss}
+\caption{FISTA Loss}
+\label{fig:}
+\end{figure}
 
@@
 
-\newpage
-\subsection*{4.3}
+We can also implement the Randomised Coordinate Projected Gradient Algorithm on the dual problem and plot the dual objective function:
 
@@
 \centering
-\includegraphics[scale=0.25]{outputs/part_4/initial-contour}
-\caption{$\alpha^{(0)}$ Contour Plot}
+\includegraphics[scale=0.3]{outputs/part_4/rcpga-loss}
+\caption{RCPGA Loss}
 \label{fig:}
 \end{figure}
+\newpage
+
+\subsection*{4.3}
+
+We can plot the decision boundary for the randomly initialised $\alpha$ vector as well as the two classes:
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.35]{outputs/part_4/initial-contour}
+\caption{Initial Contour Plot}
+\label{fig:}
+\end{figure}
+
+Similarly for each algorithm, we can plot the decision boundaries from the learned $\alpha$:
 
@@
   \centering
-\includegraphics[scale=0.25]{outputs/part_4/rcpga-contour}
+\includegraphics[scale=0.35]{outputs/part_4/rcpga-contour}
 \caption{RCPGA Contour Plot}
@@
   \centering
-\includegraphics[scale=0.25]{outputs/part_4/fista-contour}
+\includegraphics[scale=0.35]{outputs/part_4/fista-contour}
 \caption{FISTA Contour Plot}
@@
 \subsection*{4.4}
-
-\begin{figure}[h]
-\centering
-\begin{minipage}{.5\textwidth}
-  \centering
-\includegraphics[scale=0.4]{outputs/part_4/rcpga-loss}
-\caption{RCPGA Loss}
-\label{fig:}
-\end{minipage}%
-\begin{minipage}{.5\textwidth}
-  \centering
-\includegraphics[scale=0.4]{outputs/part_4/fista-loss}
-\caption{FISTA Loss}
-\label{fig:}
-\end{minipage}
-\end{figure}
+Comparing the two algorithms, we can see that both are able to converge to $\alpha$ vectors producing decision boundaries that successfully separate the two training points for the two classes.
+However, the FISTA decision boundary is quite close to the points for both classes (i.e.
+the decision boundary is not as curves as for RCPGA).
+Thus, the quality of the solution is not as nice as the RCPGA where there is ample buffer surrounding the decision boundary for both classes.
+This would seem that RCPGA can produce more reasonable results, however considering the loss function plots for both algorithms, we can see that this comes at a tradeoff of computational resources/speed.
+FISTA is able to converge in around 50 iterations whereas RCPGA required on the order of $3\times 10^{5}$ iterations.
+Although each FISTA step involves two steps (to calculate $\alpha_{k+1}$ and $\nu_{k+1}$), this is still a much faster algorithm than RCGPA.
+Thus, depending on the practical use case, if there is a requirement on the algorithm's speed, then FISTA may be more appropriate.
+On the other hand, if we have more resources available, then RCGPA may be more appropriate as it seems to produce more reasonable/accurate results.
 

### 2023-01-06 part_3.py

@@ -56,8 +56,6 @@ def plot_dimensions(
 
 def part_3():
     np.random.seed(DEFAULT_SEED)
-
-    # Part 3
     part_3_output_folder = os.path.join(OUTPUTS_FOLDER, "part_3")
     if not os.path.exists(part_3_output_folder):
         os.makedirs(part_3_output_folder)

### 2023-01-06 part_4.py

@@ -55,8 +55,6 @@ def plot_contour(
 
 def part_4():
     np.random.seed(DEFAULT_SEED)
-
-    # Part 4
     part_4_output_folder = os.path.join(OUTPUTS_FOLDER, "part_4")
     if not os.path.exists(part_4_output_folder):
         os.makedirs(part_4_output_folder)
@@ -64,8 +62,7 @@ def part_4():
     number_of_samples = 200
     x0 = np.random.randn(number_of_samples).reshape(-1, 1)
     sigma = 0.4
-    lambda_parameter = 10
-    number_of_steps = int(2e5)
+    lambda_parameter = 5e-3
     half_moons_problem = HalfMoonsProblem.generate(
         number_of_samples=number_of_samples,
         noise=0.05,
@@ -80,6 +77,7 @@ def part_4():
     )
 
     # Randomized Coordinate Projected Gradient Algorithm
+    number_of_steps = int(3e5)
     rcpga = RandomizedCoordinateProjectedGradientAlgorithm(half_moons_problem)
     x_rcpga, loss_rcpga = rcpga.run(x0.copy(), lambda_parameter, number_of_steps)
     plot_contour(

### 2023-01-06 src/problems.py

@@ -238,7 +238,7 @@ class HalfMoonsProblem(Problem):
         :param sigma: kernel length scale
         :return: kernel evaluation k(x, y)
         """
-        diff = x - y
+        diff = np.abs(x - y)
         return np.exp(-diff.T @ diff / (2 * sigma**2)).item()
 
     @staticmethod
@@ -277,14 +277,15 @@ class HalfMoonsProblem(Problem):
         return np.clip(x, 0, 1 / (lambda_parameter * self.n))
 
     @staticmethod
-    def indicator_function(x) -> float:
+    def indicator_function(x, scale) -> float:
         """
         Indicator function on [0, 1]^n
 
         :param x: current solution (number of training points, 1)
-        :return: if x lies in [0, 1]^n then 0, otherwise infinity
+        :param scale: upper bound
+        :return: if x lies in [0, scale]^n then 0, otherwise infinity
         """
-        return float("inf") if np.sum(x < 0) + np.sum(x > 1) else 0
+        return float("inf") if np.sum(x < 0) + np.sum(x > scale) else 0
 
     def loss(self, x: np.ndarray, lambda_parameter: float) -> float:
         """
@@ -297,8 +298,8 @@ class HalfMoonsProblem(Problem):
         scale = 1 / (lambda_parameter * self.n)
         return (
             0.5 * x.T @ self.a_matrix @ x
-            - scale * np.sum(x)
-            + self.indicator_function(x / scale)
+            - np.sum(x)
+            + self.indicator_function(x, scale)
         ).item()
 
     def grad_f_j(self, x: np.ndarray, j) -> float:

### 2023-01-06 submission.tex
edit report
update svm plot

@@
 We are considering the optimisation problem:
-
 \[\min_{\|\textbf{A}\textbf{x} - \textbf{b}\|_{\infty} \leq \epsilon} \frac{1}{2}\|\textbf{x}\|^2\]
@@
 To compute the dual problem, we begin by reformulating the problem as:
-
 \[\min_{\textbf{x}}  f(\textbf{x}) + g(\textbf{Ax})\]
@@
 We know from the Fenchel-Rockafellar duality theory that the dual problem is formulated as:
-
 \[\min_{\textbf{u}}  g^*(\textbf{u}) + f^*(-\textbf{A}^*\textbf{u})\]
@@
 The Fenchel conjugates are:
-
 \[g^*(\textbf{Ax}) = i^*_{[0, 1]}\left(\frac{1}{\epsilon}\|\textbf{A}\textbf{x} - \textbf{b}\|_{\infty}\right) = \frac{1}{\epsilon} \| \textbf{A}\textbf{x} - \textbf{b}\|_1\]
@@
 We can substitute:
-
 \[\min_{\textbf{u}}  \frac{1}{\epsilon} \| \textbf{u} - \textbf{b}\|_1 +  \frac{1}{2}\|-\textbf{A}^*\textbf{u}\|^2\]
@@
 Simplifying, we have our dual problem:
-
 \[\min_{\textbf{u}} \frac{1}{2}\textbf{u}^*\textbf{A}\textbf{A}^*\textbf{u} + \frac{1}{\epsilon} \| \textbf{u} - \textbf{b}\|_1\]
@@
 We know that for $f(x) = \frac{1}{2}\|\textbf{x}\|^2$:
-
 \[\textbf{0}_d \in int \big(dom(f)\big)\]
@@
 where $\textbf{0}_d = \textbf{0} \in \mathbb{R}^{d \times 1}$. Thus, knowing that $\textbf{A} \in \mathbb{R}^{n\times d}$ is a linear operator:
-
 \[\textbf{0}_n \in int \big(\textbf{A} dom(f)\big)\]
@@
 Moreover, for $g(\textbf{u}) =  i_{[0, 1]}\left(\frac{1}{\epsilon}\|\textbf{u} - \textbf{b}\|_{\infty}\right)$:
-
 \[\textbf{0}_n \in dom(g)\]
@@
 if $\|\textbf{b}\|_{\infty} \leq \epsilon$. Therefore:
-
 \[\textbf{0}_n \in int \big(dom(g)\big)\]
@@
 The KKT conditions:
-
 \[\textbf{x} \in \partial f^*(-\textbf{A}^*\textbf{u}) \text{ and } \textbf{Ax} \in \partial g^*(\textbf{u})\]
@@
 We know that:
-
 \[\partial f^*(-\textbf{A}^*\textbf{u}) = \partial \frac{1}{2}\|-\textbf{A}^*\textbf{u}\|^2 = -\textbf{A}^*\textbf{u}\]
@@
 Moreover:
-
 \[\partial g^*(\textbf{u}) = \partial\frac{1}{\epsilon} \| \textbf{u} - \textbf{b}\|_1 \]
@@
 and so
-
 \[\left(\partial\| \textbf{u} - \textbf{b}\|_1\right)_i = \begin{cases}
@@
 Thus, our KKT conditions state that:
-
 \[\textbf{x} = -\textbf{A}^*\textbf{u}\]
@@
 and
-
 \[(\textbf{Ax})_i \in \begin{cases}
@@
 To derive a rate of convergence on the primal iterates from the applications of FISTA (Fast Iterative Shrinkage Threshold Algorithm) on the  dual problem when strong duality holds, the distance to the primal solution is bounded by the dual objective values:
-
 \[\frac{2}{\mu} \|\textbf{x}_k - \bar{\textbf{x}}\|^2 \leq \Psi(\textbf{u}_k) - \Psi(\bar{\textbf{u}})\]
@@
 We also know for FISTA that (proof in lecture):
-
 \[\Psi(\textbf{u}_k) - \Psi(\bar{\textbf{u}}) \leq \frac{\|\textbf{u}_0 - \bar{\textbf{u}}\|^2}{2 \gamma \left(t_{k-1}}\right)^2\]
@@
 Thus:
-
 \[ \|\textbf{x}_k - \bar{\textbf{x}}\|^2 \leq \frac{\|\textbf{u}_0 - \bar{\textbf{u}}\|^2}{4 \gamma \left(t_{k-1}}\right)^2\]
@@
 Simplifying:
-
 \[ \|\textbf{x}_k - \bar{\textbf{x}}\| \leq \frac{\|\textbf{u}_0 - \bar{\textbf{u}}\|}{2  \sqrt{\gamma}t_{k-1}}\]
@@
 Two choices for $t_k$ in FISTA are:
-
 \begin{itemize}
@@
 In both cases the rate of convergence of the primal iterates when applying FISTA on the dual problem:
-
 \[ \|\textbf{x}_k - \bar{\textbf{x}}\| \leq \frac{a\|\textbf{u}_0 - \bar{\textbf{u}}\|}{\sqrt{\gamma}(k+1)}\]
@@
 Our problem:
-
 \[\min_{\textbf{x} \in \mathbb{R}^d} \frac{1}{2n} \|\textbf{Ax}-\textbf{y}\|^2 + \lambda \|x\|_1\]
@@
 Equivalently, our problem:
-
 \[\min_{\textbf{x} \in \mathbb{R}^d} \frac{1}{2n} \sum_{i=1}^n \left(\left\langle \textbf{a}^i, \textbf{x} \right\rangle - y_i \right)^2 + \lambda \|x\|_1\]
@@
 
-The initial vector passed to the algorithms:
+The randomly initialised vector passed to the algorithms:
 
@@
 \end{figure}
+
+We can see that both algorithms are able to recover a sparse solution, however RCPGA was able to converge in much fewer iterations compared to PSGA.
 
@@
 \subsection*{4.4}
-Comparing the two algorithms, we can see that both are able to converge to $\alpha$ vectors producing decision boundaries that successfully separate the two training points for the two classes.
-However, the FISTA decision boundary is quite close to the points for both classes (i.e.
-the decision boundary is not as curves as for RCPGA).
+Comparing the two algorithms, we can see that both are able to converge to $\alpha$ vectors producing decision boundaries that are relatively successful at separating the majority of the training points for the two classes.
+However, the FISTA decision boundary misses a couple points for both classes with a decision boundary that is rather close to boundary points of each class.
 Thus, the quality of the solution is not as nice as the RCPGA where there is ample buffer surrounding the decision boundary for both classes.
-This would seem that RCPGA can produce more reasonable results, however considering the loss function plots for both algorithms, we can see that this comes at a tradeoff of computational resources/speed.
+This would seem that RCPGA can produce better results, however considering the loss function plots for both algorithms, we can see that this comes at a tradeoff of computational resources/speed.
 FISTA is able to converge in around 50 iterations whereas RCPGA required on the order of $3\times 10^{5}$ iterations.
 Although each FISTA step involves two steps (to calculate $\alpha_{k+1}$ and $\nu_{k+1}$), this is still a much faster algorithm than RCGPA.
-Thus, depending on the practical use case, if there is a requirement on the algorithm's speed, then FISTA may be more appropriate.
+Thus, depending on the practical use case, if there is a requirement for algorithm speed, then FISTA may be more appropriate.
 On the other hand, if we have more resources available, then RCGPA may be more appropriate as it seems to produce more reasonable/accurate results.

### 2023-01-06 part_4.py

@@ -16,10 +16,10 @@ from src.problems import HalfMoonsProblem
 def plot_contour(
     half_moons_problem: HalfMoonsProblem, x: np.ndarray, algorithm: str, save_path: str
 ) -> None:
-    x_min = np.min(half_moons_problem.x[:, 0]) - 1
-    x_max = np.max(half_moons_problem.x[:, 0]) + 1
-    y_min = np.min(half_moons_problem.x[:, 1]) - 1
-    y_max = np.max(half_moons_problem.x[:, 1]) + 1
+    x_min = np.min(half_moons_problem.x[:, 0]) - 0.1
+    x_max = np.max(half_moons_problem.x[:, 0]) + 0.1
+    y_min = np.min(half_moons_problem.x[:, 1]) - 0.1
+    y_max = np.max(half_moons_problem.x[:, 1]) + 0.1
     x_ticks = np.linspace(
         x_min,
         x_max,
@@ -36,18 +36,28 @@ def plot_contour(
 
     y_grid = half_moons_problem.predict(x.T, x_grid).reshape(-1)
     y = half_moons_problem.y.reshape(-1)
-    fig = plt.figure()
-    fig.set_figwidth(2 * (x_max - x_min))
-    fig.set_figwidth(2 * (y_max - y_min))
-    plt.contourf(x1_grid, x2_grid, y_grid.reshape(x1_grid.shape))
+    fig, ax = plt.subplots(constrained_layout=True)
+    fig.set_figwidth(3 * (x_max - x_min))
+    fig.set_figheight(3 * (y_max - y_min))
+    cmap = plt.get_cmap("binary_r", 2)
+    contour = ax.contourf(x1_grid, x2_grid, y_grid.reshape(x1_grid.shape), cmap=cmap)
     for label in set(y):
+        edge = "gray"
+        if label == -1:
+            colour = "black"
+        else:
+            colour = "white"
         idx = y == label
         plt.scatter(
             half_moons_problem.x[idx, 0],
             half_moons_problem.x[idx, 1],
             label=label,
+            c=colour,
+            edgecolors=edge,
+            s=120,
         )
     plt.legend()
+    fig.colorbar(contour, ticks=[-1, 1])
     plt.title(f"Contour Plot ({algorithm})")
     plt.savefig(save_path, bbox_inches="tight")
     plt.close()

### 2023-01-06 part_4.py
update figures

@@ -23,12 +23,12 @@ def plot_contour(
     x_ticks = np.linspace(
         x_min,
         x_max,
-        100,
+        500,
     )
     y_ticks = np.linspace(
         y_min,
         y_max,
-        100,
+        500,
     )
 
     x1_grid, x2_grid = np.meshgrid(x_ticks, y_ticks)

### 2023-01-07 submission.tex
add report edits

@@
 The function for (a):
-
 \[\max\{ax+b, x^4-5, e^{x^2}\}\]
@@
 For the function:
-
 \[f(x) = \begin{cases}
-      -x \tab \text{if } x \in ]-1, 0] \\
-      x^2 \tab \text{if } x \geq 0 \\
+      -x, & \text{if } x \in ]-1, 0] \\
+      x^2, & \text{if } x \geq 0 \\
    \end{cases}\]
@@
 the sub-differential is:
-
 \[\partial f(x) = \begin{cases}
-      -1 \tab \text{if } x \in ]-1, 0[ \\
-      \left[0, 1\right] \tab \text{if } x = 0  \\
-      2x \tab \text{if } x > 0 \\
+      -1, & \text{if } x \in ]-1, 0[ \\
+      \left[0, 1\right], & \text{if } x = 0  \\
+      2x, & \text{if } x > 0 \\
    \end{cases}\]
@@
 \subsection*{1.3}
-For a function:
-\[f(x) = \langle Ax, x \rangle + \langle x, b\rangle + c\]
-
-where A is a square matrix not necessarily symmetric, the gradient is (a):
-
+For a function $f(x) = \langle Ax, x \rangle + \langle x, b\rangle + c$ where A is a square matrix not necessarily symmetric, the gradient is (a):
 \[\nabla f(x) = A^*x + Ax + b\]
-
 \subsection*{1.4}
-
 The Fenchel conjugate of $f(x) = g(2x)$ is (a):
-
 \[f^*(u) = g^*(u/2)\]
@@
 \subsection*{1.5}
-
 The solution to the dual problem is (c):
-
 \[\bar{u} = (\textbf{K}+\lambda n \textbf{Id})^{-1} y\]
 
+\newpage
 \section*{Part 2: Theory on convex analysis and optimization}
 \subsection*{2.1}
+
+The Fenchel conjugate is defined as:
+
+\[f^*(u) = \sup_{x\in \mathcal{X}} \{ \langle x, u\rangle - f(x) \}\]
 
@@
 \[ f(x) = \begin{cases}
-      +\infty \tab \text{if } x \leq 0 \\
-      -\log x \tab \text{if } x > 0 \\
+      +\infty, & \text{if } x \leq 0 \\
+      -\log x, & \text{if } x > 0 \\
    \end{cases}\]
 
-The Fenchel conjugate is defined:
+Substituting into the Fenchel conjugate definition:
 
@@
 
-To find the supremum, we can take the partial derivative with respect to $x$, set to zero, and solve for x:
+To find the supremum, we can take the partial derivative with respect to $x$, set to zero, and solve for $x$:
 
@@
 
-Thus, the supremum above is solved when $x=\frac{-1}{u}$:
+Thus, the supremum above is acheived when $x=\frac{-1}{u}$:
 
@@
 
-\[f^*(u) = - (1 + \log u)\]
+\[ f(x) = \begin{cases}
+      - (1 + \log (-u), & \text{if } u < 0 \\
+      +\infty, & \text{if } u \geq 0 \\
+   \end{cases}\]
 
@@
 
-The Fenchel conjugate is defined:
+Substituting into the Fenchel conjugate definition:
 
@@
 
-We can compute the partial derivative:
+We can compute the partial derivative with respect to $x$, set to zero, and solve for $x$:
 
@@
 
-Thus, the supremum is solved when $x=\frac{u}{2}$:
+Thus, the supremum is achieved when $x=\frac{u}{2}$:
 
@@
 
-We can see the Fenchel conjugate is:
+We can see that the Fenchel conjugate is:
 
 \[ f^*(u) = \max(0, u)}\]
+
+is the ReLu function.
 
@@
 
-for all $x1, x2 \in \mathcal{X}$ and $\lambda_1, \lambda_2 \in \mathbb{R}_+$ with $\lambda_1 + \lambda_2 = 1$. In this base case $n=2$.
-
-Our inductive step will prove that the inequality continues to hold for $n+1$:
+for all $x_1, x_2 \in \mathcal{X}$ and $\lambda_1, \lambda_2 \in \mathbb{R}_+$ with $\lambda_1 + \lambda_2 = 1$.
+In this base case $n=2$.
+Our inductive step will prove that the inequality continues to hold for $n+1$.
+We begin with:
 
@@
 
-We can insert the term $(1-\lambda_{n+1})$:
+where $\sum_{i=1}^{n+1} \lambda_i = 1$. We can insert the term $(1-\lambda_{n+1})$:
 
@@
 
-Rewriting this,
+Substituting the definition for $\bar{x}$:
 
@@
 
-We know that the first term on the right hand side:
+We know that from the known inequality from the $n$ case:
 
@@
 
-Thus can upper bound our previous inequality:
+Thus we can upper bound our previous inequality:
 
-\[f\left(\sum_{i=1}^{n+1} \lambda_i x_i\right)  \leq \sum_{i=1}^n \frac{\lambda_i}{{1-\lambda_{n+1}}}  f \left(x_i\right) + \lambda_{n+1} f\left( x_{n+1}\right) =\sum_{i=1}^{n+1} \frac{\lambda_i}{{1-\lambda_{n+1}}} f \left(x_i\right) \]
+\[f\left(\sum_{i=1}^{n+1} \lambda_i x_i\right)  \leq (1-\lambda_{n+1})\sum_{i=1}^n \frac{\lambda_i}{{1-\lambda_{n+1}}}  f \left(x_i\right) + \lambda_{n+1} f\left( x_{n+1}\right) =\sum_{i=1}^{n+1} \lambda_i f \left(x_i\right) \]
 
-proving our inductive step and Jensen's inequality as required. \square
+proving our inductive step for $n+1$ and Jensen's inequality as required. \square
 
@@
 
-\[\text{f is convex} \leftrightarrow \langle \nabla f(x) - \nabla f(y), x-y \rangle \geq 0\]
+\[\text{A function } f \text{ is convex} \leftrightarrow \langle \nabla f(x) - \nabla f(y), x-y \rangle \geq 0\]
 
@@
 \begin{gather*}
-    \frac{1}{y} > \frac{1}{x}\\
     x-y>0\\
@@
 \begin{gather*}
-    \frac{1}{y} < \frac{1}{x}\\
     x-y<0\\
@@
 \begin{gather*}
-    \frac{1}{y} = \frac{1}{x}\\
     x-y=0\\
@@
 
-Thus, $\forall x, y \in dom f, \langle \nabla f(x) - \nabla f(y), x-y \rangle \geq 0$ so $f(x) = -\log(x)$ is convex. \square
+Thus, $\forall x, y \in dom f, \langle \nabla f(x) - \nabla f(y), x-y \rangle \geq 0$ so by the characterization for differentiable functions, $f(x) = -\log(x)$ is convex. \square
 
@@
 
-\[\frac{1}{n}\sum_{i=1}^n x_i \geq \sqrt[n]{x_1 \cdots x_n}\] \square
+\[\frac{1}{n}\sum_{i=1}^n x_i \geq \sqrt[n]{x_1 \cdots x_n}\]
+
+for all $x_1, ..., x_n \in \mathbb{R}^+$ \square
+
 \subsection*{2.3}
@@
 
-where $\sum_{i=1}^{m} \lambda_i = 1$. Knowing that $f$ is a convex function on $C$, we can write Jensen's inequality for :
+where $\sum_{i=1}^{m} \lambda_i = 1$. Knowing that $f$ is a convex function on $C$, we can write Jensen's inequality:
 
@@
 
-where $\sum_{i=1}^{m} \lambda_i = 1$.
-
 Because $\sum_{i=1}^{m} \lambda_i f(a_i)$ is a weighted average of $f(a_i)$'s we know that:
 
-\[\sum_{i=1}^{m} \lambda_i f(a_i) \leq \max_{\{a_i\}_{i=1}^m} f(a_i)\]
+\[\sum_{i=1}^{m} \lambda_i f(a_i) \leq \max_{i} f(a_i)\]
 
@@
 
-\[f\left(\sum_{i=1}^{m} \lambda_i a_i \right) \leq \max_{\{a_i\}_{i=1}^m} f(a_i)\]
+\[f\left(\sum_{i=1}^{m} \lambda_i a_i \right) \leq \max_{i} f(a_i)\]
 
-So the maximum of the convex function $f$ on $C$ is attained at one of the vertices $a_1, ..., a_m$.
+So all function evaluations of $f(x)$ for $x \in C$ are less than or equal to the maximum of the function evaluations at the verticies $a_1, ..., a_m$ and so the maximum of the convex function $f$ on $C$ is attained at one of the vertices $a_1, ..., a_m$.
 \square
@@
 \[\nabla^2 f(x, y) = \begin{bmatrix}
-                 2 & 4\\
-                 4 & 8
+                 2 & -4\\
+                 -4 & 8
          \end{bmatrix}
@@
 
-we see that the Hessian is positive semi-definite and so the function is jointly convex.
+we have our Hessian matrix for which the determinant is $(2)(8) - (-4)^2 \geq 0$ so we see that the Hessian is positive semi-definite and so the function is jointly convex.
 
@@
   The conditions for the uniqueness of minimizers is that $f$ is strictly convex.
-Thus for the existence and uniqueness of minimizers for a convex function $f$, we require that $f$ is closed, coercive, and strictly convex.
+Thus the minimal sufficient conditions for the existence and uniqueness of minimizers for a convex function $f$, are that $f$ is closed, coercive, and strictly convex.
 
@@
 
-where $f(x) = \frac{1}{2}\|\textbf{x}\|^2$ and $g(\textbf{Ax}) =  i_{[0, 1]}\left(\frac{1}{\epsilon}\|\textbf{A}\textbf{x} - \textbf{b}\|_{\infty}\right)$.
+where $f(\textbf{x}) = \frac{1}{2}\|\textbf{x}\|^2$ and $g(\textbf{Ax}) =  i_{\frac{1}{\epsilon}\|\textbf{A}\textbf{x} - \textbf{b}\|_{\infty} \leq 1}\left(\textbf{Ax}\right)$, the indicator function of the unit ball for $\frac{1}{\epsilon}\|\textbf{A}\textbf{x} - \textbf{b}\|_{\infty}$.
 
@@
 
-\[\min_{\textbf{x}}  \frac{1}{2}\|\textbf{x}\|^2 + i_{[0, 1]}\left(\frac{1}{\epsilon}\|\textbf{A}\textbf{x} - \textbf{b}\|_{\infty}\right)\]
+\[\min_{\textbf{x}}  \frac{1}{2}\|\textbf{x}\|^2 + i_{\frac{1}{\epsilon}\|\textbf{A}\textbf{x} - \textbf{b}\|_{\infty} \leq 1}\left(\textbf{Ax}\right)\]
 
@@
 
-The Fenchel conjugates are:
-\[g^*(\textbf{Ax}) = i^*_{[0, 1]}\left(\frac{1}{\epsilon}\|\textbf{A}\textbf{x} - \textbf{b}\|_{\infty}\right) = \frac{1}{\epsilon} \| \textbf{A}\textbf{x} - \textbf{b}\|_1\]
+For the Fenchel conjugate of $g(\textbf{Ax})$, we first recognize that the dual of the indicator function is the support function under $\frac{1}{\epsilon}\|\textbf{A}\textbf{x} - \textbf{b}\|_{\infty}$ which is the definition for the dual norm $\frac{1}{\epsilon}\|\textbf{A}\textbf{x} - \textbf{b}\|_{1}$.
+
+Thus:
+
+\[g^*(\textbf{Ax}) = i^*_{\frac{1}{\epsilon}\|\textbf{A}\textbf{x} - \textbf{b}\|_{\infty} \leq 1}\left(\textbf{Ax}\right) = \frac{1}{\epsilon} \| \textbf{A}\textbf{x} - \textbf{b}\|_1\].
+
+The Fenchel conjugate of $f(\textbf{x})$:
 
@@
 
-Moreover, for $g(\textbf{u}) =  i_{[0, 1]}\left(\frac{1}{\epsilon}\|\textbf{u} - \textbf{b}\|_{\infty}\right)$:
+Moreover, for $g(\textbf{u}) =  i_{\frac{1}{\epsilon}\|\textbf{u} - \textbf{b}\|_{\infty} \leq 1}\left(\textbf{u}\right)$:
 \[\textbf{0}_n \in dom(g)\]
 
-if $\|\textbf{b}\|_{\infty} \leq \epsilon$. Therefore:
+if $\|\textbf{b}\|_{\infty} \leq \epsilon$ and,
 \[\textbf{0}_n \in int \big(dom(g)\big)\]
@@
 
-where $\partial f^*$ and $\partial g^*$ are the subgradients of $f^*$ and $g^*$ respectively.
+and equivalently,
+
+\[-\textbf{A}^*\textbf{u} \in \partial f(\textbf{x}) \text{ and } \textbf{u} \in \partial g(\textbf{Ax})\]
+
+we will use the formulation involving $\partial f^*$ and $\partial g^*$, the subgradients of $f^*$ and $g^*$ respectively.
 
@@
 and so
-\[\left(\partial\| \textbf{u} - \textbf{b}\|_1\right)_i = \begin{cases}
+\[\left(\partial \frac{1}{\epsilon}\| \textbf{u} - \textbf{b}\|_1\right)_i = \begin{cases}
       \frac{1}{\epsilon}, &  \text{if } | u_i - b_i| > 0 \\
@@
 
-where $\left(\partial\| \textbf{u} - \textbf{b}\|_1\right)_i$ is the $i^{th}$ element of $\partial\| \textbf{u} - \textbf{b}\|_1 \in \mathbb{R}^{n \times 1}$
+where $\left(\partial\| \textbf{u} - \textbf{b}\|_1\right)_i$ is the $i^{th}$ element of $\partial\| \textbf{u} - \textbf{b}\|_1 \in \mathbb{R}^{n \times 1}$, for $i=1, ..., n$.
 
@@
 However, the FISTA decision boundary misses a couple points for both classes with a decision boundary that is rather close to boundary points of each class.
-Thus, the quality of the solution is not as nice as the RCPGA where there is ample buffer surrounding the decision boundary for both classes.
+Thus, the quality of the solution is not as nice as the RCPGA where there are ample buffer zones surrounding the decision boundary for both classes.
 This would seem that RCPGA can produce better results, however considering the loss function plots for both algorithms, we can see that this comes at a tradeoff of computational resources/speed.

### 2023-01-07 part_4.py

@@ -8,7 +8,7 @@ from constants import DEFAULT_SEED, OUTPUTS_FOLDER
 from helpers import plot_loss
 from src.gradient_algorithms import (
     FastIterativeShrinkageThresholdAlgorithm,
-    RandomizedCoordinateProjectedGradientAlgorithm,
+    RandomizedCoordinateProximalGradientAlgorithm,
 )
 from src.problems import HalfMoonsProblem
 
@@ -23,12 +23,12 @@ def plot_contour(
     x_ticks = np.linspace(
         x_min,
         x_max,
-        500,
+        1000,
     )
     y_ticks = np.linspace(
         y_min,
         y_max,
-        500,
+        1000,
     )
 
     x1_grid, x2_grid = np.meshgrid(x_ticks, y_ticks)
@@ -88,7 +88,7 @@ def part_4():
 
     # Randomized Coordinate Projected Gradient Algorithm
     number_of_steps = int(3e5)
-    rcpga = RandomizedCoordinateProjectedGradientAlgorithm(half_moons_problem)
+    rcpga = RandomizedCoordinateProximalGradientAlgorithm(half_moons_problem)
     x_rcpga, loss_rcpga = rcpga.run(x0.copy(), lambda_parameter, number_of_steps)
     plot_contour(
         half_moons_problem=half_moons_problem,

### 2023-01-07 src/gradient_algorithms.py

@@ -107,7 +107,7 @@ class RandomizedCoordinateProximalGradientAlgorithm(GradientAlgorithm):
         :param j: dimension index
         :return: gamma_j for the rcpga algorithm
         """
-        return self.problem.n / (np.linalg.norm(self.problem.a_matrix_j(j)) ** 2)
+        return self.problem.mu / (np.linalg.norm(self.problem.a_matrix_j(j)) ** 2)
 
     def run(
         self, x, lambda_parameter, number_of_steps
@@ -134,22 +134,6 @@ class RandomizedCoordinateProximalGradientAlgorithm(GradientAlgorithm):
             )
             loss.append(self.problem.loss(x, lambda_parameter))
         return x, loss
-
-
-class RandomizedCoordinateProjectedGradientAlgorithm(
-    RandomizedCoordinateProximalGradientAlgorithm
-):
-    def __init__(self, problem: HalfMoonsProblem):
-        super().__init__(problem)
-
-    def calculate_gamma_parameter(self, j: int) -> float:
-        """
-        gamma = 1/(||a_j||^2)
-
-        :param j: dimension index
-        :return: gamma_j for the rcpga algorithm
-        """
-        return 1 / (np.linalg.norm(self.problem.a_matrix_j(j)) ** 2)
 
 
 class FastIterativeShrinkageThresholdAlgorithm(GradientAlgorithm):

### 2023-01-07 src/problems.py

@@ -104,6 +104,14 @@ class Problem(ABC):
         """
         pass
 
+    @property
+    @abstractmethod
+    def mu(self) -> float:
+        """
+        :return: strong convexity coefficient
+        """
+        pass
+
 
 class SparseProblem(Problem):
     def __init__(
@@ -129,6 +137,10 @@ class SparseProblem(Problem):
         self.sparsity = sparsity
         self.std = std
         super().__init__(a_matrix, y)
+
+    @property
+    def mu(self) -> float:
+        return self.n
 
     @staticmethod
     def proximity_operator(
@@ -227,6 +239,10 @@ class HalfMoonsProblem(Problem):
             a_matrix=np.multiply(y @ y.T, self.gram_matrix),
             y=y,
         )
+
+    @property
+    def mu(self) -> float:
+        return 1
 
     @staticmethod
     def kernel(x: np.ndarray, y: np.ndarray, sigma: float) -> float:

### 2023-01-07 src/gradient_algorithms.py
update docstring

@@ -37,13 +37,13 @@ class ProximalStochasticGradientAlgorithm(GradientAlgorithm):
 
     def calculate_gamma_parameter(self, k: int):
         """
-        gamma_k = n/(||A||^2 * sqrt (k+1))
+        gamma_k = mu/(||A||^2 * sqrt (k+1))
 
         :param k: step number
         :return: gamma_k for the psga algorithm
         """
 
-        return self.problem.n / (
+        return self.problem.mu / (
             np.linalg.norm(self.problem.a_matrix, ord=np.inf) ** 2 * np.sqrt(k + 1)
         )
 
@@ -102,7 +102,7 @@ class RandomizedCoordinateProximalGradientAlgorithm(GradientAlgorithm):
 
     def calculate_gamma_parameter(self, j: int) -> float:
         """
-        gamma = n/(||a_j||^2)
+        gamma = mu/(||a_j||^2)
 
         :param j: dimension index
         :return: gamma_j for the rcpga algorithm

### 2023-01-07 src/gradient_algorithms.py
update docstrings

@@ -14,6 +14,14 @@ class GradientAlgorithm(ABC):
     def run(
         self, x: np.ndarray, lambda_parameter: float, number_of_steps: int
     ) -> Tuple[np.ndarray, List[float]]:
+        """
+        Run the gradient based algorithm.
+
+        :param x: initial solution (number of dimension, 1)
+        :param lambda_parameter: scalar on g(x) of the minimization problem
+        :param number_of_steps: number of steps to run the algorithm
+        :return: optimised solution from algorithm and list of loss values at each iteration
+        """
         pass
 
 
@@ -22,7 +30,7 @@ class ProximalStochasticGradientAlgorithm(GradientAlgorithm):
         """
         The Proximal Stochastic Gradient Algorithm.
 
-        :param problem: problem object
+        :param problem: sparse problem
         :param is_ergodic_mean: whether to use the ergodic mean for loss calculation
         """
         self.is_ergodic_mean = is_ergodic_mean
@@ -91,6 +99,11 @@ class ProximalStochasticGradientAlgorithm(GradientAlgorithm):
 
 class RandomizedCoordinateProximalGradientAlgorithm(GradientAlgorithm):
     def __init__(self, problem: Problem):
+        """
+        The Randomized Coordinate Proximal Gradient Algorithm.
+
+        :param problem: the problem
+        """
         super().__init__(problem)
 
     def generate_random_dimension_index(self) -> int:
@@ -103,6 +116,7 @@ class RandomizedCoordinateProximalGradientAlgorithm(GradientAlgorithm):
     def calculate_gamma_parameter(self, j: int) -> float:
         """
         gamma = mu/(||a_j||^2)
+        where mu is the coefficient of strong convexity
 
         :param j: dimension index
         :return: gamma_j for the rcpga algorithm
@@ -138,6 +152,11 @@ class RandomizedCoordinateProximalGradientAlgorithm(GradientAlgorithm):
 
 class FastIterativeShrinkageThresholdAlgorithm(GradientAlgorithm):
     def __init__(self, problem: HalfMoonsProblem):
+        """
+        The Fast Iterative Shrinkage Threshold Algorithm.
+
+        :param problem: half moons problem
+        """
         super().__init__(problem)
         self.problem = problem
 

### 2023-01-07 submission.tex
update plots and report

@@
 \centering
-\includegraphics[scale=0.35]{outputs/part_4/initial-contour}
+\includegraphics[scale=0.3]{outputs/part_4/initial-contour}
 \caption{Initial Contour Plot}
@@
   \centering
-\includegraphics[scale=0.35]{outputs/part_4/rcpga-contour}
+\includegraphics[scale=0.3]{outputs/part_4/rcpga-contour}
 \caption{RCPGA Contour Plot}
@@
   \centering
-\includegraphics[scale=0.35]{outputs/part_4/fista-contour}
+\includegraphics[scale=0.3]{outputs/part_4/fista-contour}
 \caption{FISTA Contour Plot}
@@
 \subsection*{4.4}
-Comparing the two algorithms, we can see that both are able to converge to $\alpha$ vectors producing decision boundaries that are relatively successful at separating the majority of the training points for the two classes.
-However, the FISTA decision boundary misses a couple points for both classes with a decision boundary that is rather close to boundary points of each class.
-Thus, the quality of the solution is not as nice as the RCPGA where there are ample buffer zones surrounding the decision boundary for both classes.
-This would seem that RCPGA can produce better results, however considering the loss function plots for both algorithms, we can see that this comes at a tradeoff of computational resources/speed.
-FISTA is able to converge in around 50 iterations whereas RCPGA required on the order of $3\times 10^{5}$ iterations.
+Comparing the two algorithms, we can see that both are able to converge to $\alpha$ vectors producing decision boundaries that are successful at separating the training points for the two classes.
+It is difficult to determine which of the two decision boundaries are more suitable for the training data, without more context about the data itself.
+However considering the loss function plots for both algorithms, we can see that FISTA is able to converge in around 20 iterations whereas RCPGA required on the order of $3\times 10^{5}$ iterations.
 Although each FISTA step involves two steps (to calculate $\alpha_{k+1}$ and $\nu_{k+1}$), this is still a much faster algorithm than RCGPA.
-Thus, depending on the practical use case, if there is a requirement for algorithm speed, then FISTA may be more appropriate.
-On the other hand, if we have more resources available, then RCGPA may be more appropriate as it seems to produce more reasonable/accurate results.
+However, we can also see that the objective function for FISTA is much less smooth compared to RCGPA.
+Depending on use case, this may or may not be a problem.
+Overall, FISTA is able to produce quality decision boundaries much faster than RCGPA.
 

### 2023-01-07 part_3.py

@@ -55,7 +55,6 @@ def plot_dimensions(
 
 
 def part_3():
-    np.random.seed(DEFAULT_SEED)
     part_3_output_folder = os.path.join(OUTPUTS_FOLDER, "part_3")
     if not os.path.exists(part_3_output_folder):
         os.makedirs(part_3_output_folder)
@@ -73,10 +72,10 @@ def part_3():
         algorithm=f"Initialised Vector (x0)",
         save_path=os.path.join(part_3_output_folder, f"initial-x"),
     )
-    lambda_parameter = 3e-1
-    number_of_steps = int(5e5)
 
     # Proximal Stochastic Gradient Algorithm
+    lambda_parameter = 3e-1
+    number_of_steps = int(5e5)
     proximal_stochastic_gradient_algorithm = ProximalStochasticGradientAlgorithm(
         sparse_problem,
         is_ergodic_mean=False,
@@ -136,4 +135,5 @@ def part_3():
 
 
 if __name__ == "__main__":
+    np.random.seed(DEFAULT_SEED)
     part_3()

### 2023-01-07 part_4.py

@@ -1,5 +1,4 @@
 import os
-from typing import List
 
 import matplotlib.pyplot as plt
 import numpy as np
@@ -14,7 +13,11 @@ from src.problems import HalfMoonsProblem
 
 
 def plot_contour(
-    half_moons_problem: HalfMoonsProblem, x: np.ndarray, algorithm: str, save_path: str
+    half_moons_problem: HalfMoonsProblem,
+    x: np.ndarray,
+    contour_resolution: int,
+    algorithm: str,
+    save_path: str,
 ) -> None:
     x_min = np.min(half_moons_problem.x[:, 0]) - 0.1
     x_max = np.max(half_moons_problem.x[:, 0]) + 0.1
@@ -23,12 +26,12 @@ def plot_contour(
     x_ticks = np.linspace(
         x_min,
         x_max,
-        1000,
+        contour_resolution,
     )
     y_ticks = np.linspace(
         y_min,
         y_max,
-        1000,
+        contour_resolution,
     )
 
     x1_grid, x2_grid = np.meshgrid(x_ticks, y_ticks)
@@ -37,8 +40,8 @@ def plot_contour(
     y_grid = half_moons_problem.predict(x.T, x_grid).reshape(-1)
     y = half_moons_problem.y.reshape(-1)
     fig, ax = plt.subplots(constrained_layout=True)
-    fig.set_figwidth(3 * (x_max - x_min))
-    fig.set_figheight(3 * (y_max - y_min))
+    fig.set_figwidth(3.5 * (x_max - x_min))
+    fig.set_figheight(3.5 * (y_max - y_min))
     cmap = plt.get_cmap("binary_r", 2)
     contour = ax.contourf(x1_grid, x2_grid, y_grid.reshape(x1_grid.shape), cmap=cmap)
     for label in set(y):
@@ -54,7 +57,7 @@ def plot_contour(
             label=label,
             c=colour,
             edgecolors=edge,
-            s=120,
+            s=50,
         )
     plt.legend()
     fig.colorbar(contour, ticks=[-1, 1])
@@ -63,16 +66,18 @@ def plot_contour(
     plt.close()
 
 
-def part_4():
-    np.random.seed(DEFAULT_SEED)
+def part_4(
+    contour_resolution: int,
+):
     part_4_output_folder = os.path.join(OUTPUTS_FOLDER, "part_4")
     if not os.path.exists(part_4_output_folder):
         os.makedirs(part_4_output_folder)
 
     number_of_samples = 200
+    sigma = 0.275
[…]

### 2023-01-07 src/gradient_algorithms.py

@@ -171,11 +171,15 @@ class FastIterativeShrinkageThresholdAlgorithm(GradientAlgorithm):
 
     def calculate_gamma_parameter(self) -> float:
         """
-        gamma = 1/(||A||^2)
+        gamma = mu/L
+        where L is the lipschitz constant of A^*A
 
         :return: gamma value
         """
-        return 1 / (np.linalg.norm(self.problem.a_matrix, ord=np.inf) ** 2)
+        lipschitz_constant = np.linalg.eigvalsh(
+            self.problem.a_matrix.T @ self.problem.a_matrix
+        )[-1]
+        return self.problem.mu / lipschitz_constant
 
     def run(
         self, x, lambda_parameter, number_of_steps

### 2023-01-07 src/problems.py

@@ -28,6 +28,12 @@ class Problem(ABC):
         #  number of dimensions
         return self.a_matrix.shape[1]
 
+    @property
+    @abstractmethod
+    def mu(self) -> float:
+        # Strong convexity coefficient
+        pass
+
     def a_matrix_i(self, i: int):
         """
         ith row of the A matrix
@@ -104,14 +110,6 @@ class Problem(ABC):
         """
         pass
 
-    @property
-    @abstractmethod
-    def mu(self) -> float:
-        """
-        :return: strong convexity coefficient
-        """
-        pass
-
 
 class SparseProblem(Problem):
     def __init__(
@@ -140,6 +138,7 @@ class SparseProblem(Problem):
 
     @property
     def mu(self) -> float:
+        # Strong convexity coefficient
         return self.n
 
     @staticmethod
@@ -202,7 +201,6 @@ class SparseProblem(Problem):
         :param std: standard deviation parameterizing epsilon, the noise added to the observed response y
         :return: spare problem
         """
-
         assert sparsity % 2 == 0, f"{sparsity=} needs to be divisible by 2"
         xsp = 0.5 * (np.random.rand(sparsity // 2) + 1)
         xsn = -0.5 * (np.random.rand(sparsity // 2) + 1)
@@ -242,6 +240,7 @@ class HalfMoonsProblem(Problem):
 
     @property
     def mu(self) -> float:
+        # Strong convexity coefficient
         return 1
 
     @staticmethod
@@ -343,15 +342,14 @@ class HalfMoonsProblem(Problem):
         number_of_samples: int, noise: float, sigma: float, random_state: int
     ) -> HalfMoonsProblem:
         """
-        Generate a Half Moons Probelm
+        Generate a Half Moons Problem
 
         :param number_of_samples: number of samples to generate
         :param noise: noise in data
         :param sigma: kernel length scale
-        :param random_state: randomisation statea
+        :param random_state: randomisation state
         :return: Half Moons Problem
         """
-
         x, y = make_moons(
             n_samples=number_of_samples, noise=noise, random_state=random_state
         )

### 2023-01-08 submission.tex
edit report

@@
 
-\title{\textbf{COMP0083 Convex Optimisation Assignment}}
+\title{\textbf{COMP0083 Convex Optimisation Coursework}}
 %\author{Jian Shu (James) Wu \\ }
@@
 \subsection*{1.1}
-
-The function for (a):
+The function for (a) is a convex function:
 \[\max\{ax+b, x^4-5, e^{x^2}\}\]
-
-is a convex function.
 
@@
 \[\nabla f(x) = A^*x + Ax + b\]
+
+where $A^*$ denotes the transpose of $A$.
 \subsection*{1.4}
@@
 \subsection*{1.5}
-The solution to the dual problem is (c):
+The solution to the dual problem of the ridge regression problem is (c):
 \[\bar{u} = (\textbf{K}+\lambda n \textbf{Id})^{-1} y\]
@@
 
-The Fenchel conjugate is defined as:
-
+The Fenchel conjugate of $f(x)$ is defined as:
 \[f^*(u) = \sup_{x\in \mathcal{X}} \{ \langle x, u\rangle - f(x) \}\]
@@
 Substituting into the Fenchel conjugate definition:
-
-\[f^*(u) = \sup_{x\in \mathcal{X}} \{ \langle x, u\rangle +  \log x \}\]
+\[f^*(u) = \sup_{x > 0} \{ \langle x, u\rangle +  \log x \}\]
 
 To find the supremum, we can take the partial derivative with respect to $x$, set to zero, and solve for $x$:
-
-\[\frac{\partial}{\partial x}(ux + logx) = u + \frac{1}{x} = 0\]
+\[\frac{\partial}{\partial x}(ux + \log x) = u + \frac{1}{x} = 0\]
 
 Thus, the supremum above is acheived when $x=\frac{-1}{u}$:
-
 \[f^*(u) = u \left(\frac{-1}{u} \right) +  \log\left(\frac{-1}{u} \right) \]
 
+where $x>0$ so $u < 0$.
 Simplifying, we have the Fenchel conjugate:
-
 \[ f(x) = \begin{cases}
@@
 \subsection*{2.1.2} Given:
-
 \[f(x) = x^2\]
@@
 Substituting into the Fenchel conjugate definition:
-
 \[f^*(u) = \sup_{x\in \mathcal{X}} \{ \langle x, u\rangle - x^2 \}\]
@@
 We can compute the partial derivative with respect to $x$, set to zero, and solve for $x$:
-
 \[\frac{\partial}{\partial x}(ux - x^2) = u - 2x = 0\]
@@
 Thus, the supremum is achieved when $x=\frac{u}{2}$:
-
 \[f^*(u) = u \left(\frac{u}{2} \right) -  \left(\frac{u}{2} \right)^2 \]
@@
 Simplifying, we have the Fenchel conjugate:
-
 \[f^*(u) =  \frac{u^2}{4}\]
@@
 \subsection*{2.1.3}Given:
-
 \[f(x) = i_{\left[0, 1\right]}\]
@@
 Substituting into the Fenchel conjugate definition:
-
 \[f^*(u) = \sup_{x\in \mathcal{X}} \{ \langle x, u\rangle - i_{\left[0, 1\right]} \}\]
@@
 Thus,
-
 \[f^*(u) = \sup_{x\in \left[0, 1\right]} \{ \langle x, u\rangle \}\]
@@
 
-is the ReLu function.
+the ReLu function.
 
@@
 Given $f$ a proper convex function, to prove by induction Jensen's inequality:
-
 \[f\left(\sum_{i=1}^n \lambda_i x_i\right) \leq \sum_{i=1}^{n} \lambda_i f(x_i)\]
@@
 for all $x_1, ... x_n \in \mathcal{X}$ and for all $\lambda_1, ..., \lambda_n \in \mathbb{R}_+$ with $\sum_{i=1}^n \lambda_i = 1$, we start with the definition of convexity which states:
-
 \[f(\lambda_1 x_1 + \lambda_2 x_2) \leq \lambda_1 f(x_1) + \lambda_2 f(x_2)\]
@@
 for all $x_1, x_2 \in \mathcal{X}$ and $\lambda_1, \lambda_2 \in \mathbb{R}_+$ with $\lambda_1 + \lambda_2 = 1$.
-In this base case $n=2$.
+This acts as a base case for Jensen's inequality when $n=2$.
 Our inductive step will prove that the inequality continues to hold for $n+1$.
 We begin with:
-
 \[f\left(\sum_{i=1}^{n+1} \lambda_i x_i\right) = f\left(\sum_{i=1}^n \lambda_i x_i + \lambda_{n+1} x_{n+1}\right)\]
@@
 where $\sum_{i=1}^{n+1} \lambda_i = 1$. We can insert the term $(1-\lambda_{n+1})$:
-
 \[f\left(\sum_{i=1}^{n+1} \lambda_i x_i\right) = f\left((1-\lambda_{n+1}) \sum_{i=1}^n \frac{\lambda_i}{1-\lambda_{n+1}} x_i + \lambda_{n+1} x_{n+1}\right)\]
 
-If we define $\bar{x} = \sum_{i=1}^n \frac{\lambda_i}{1-\lambda_{n+1}} x_i$ and we are back to our $n=2$ base case, so we know from convexity:
-
+If we define $\bar{x} = \sum_{i=1}^n \frac{\lambda_i}{1-\lambda_{n+1}} x_i$, we are back to our $n=2$ base case, so we know that:
 \[ f\left((1-\lambda_{n+1})\bar{x} + \lambda_{n+1} x_{n+1}\right) \leq (1-\lambda_{n+1}) f\left( \bar{x}\right) + \lambda_{n+1} f\left( x_{n+1}\right)\]
 
-Substituting the definition for $\bar{x}$:
-
+Substituting back the definition for $\bar{x}$:
 \[f\left(\sum_{i=1}^{n+1} \lambda_i x_i\right) \leq (1-\lambda_{n+1}) f\left( \sum_{i=1}^n \frac{\lambda_i}{1-\lambda_{n+1}} x_i\right) + \lambda_{n+1} f\left( x_{n+1}\right)\]
 
-We know that from the known inequality from the $n$ case:
-
+We know that from the inequality in the $n$ case:
 \[f\left( \sum_{i=1}^n \frac{\lambda_i}{1-\lambda_{n+1}} x_i\right) = \frac{1}{{1-\lambda_{n+1}}}f\left( \sum_{i=1}^n \lambda_i x_i\right) \leq \frac{1}{{1-\lambda_{n+1}}} \sum_{i=1}^n \lambda_i f \left(x_i\right) =  \sum_{i=1}^n \frac{\lambda_i}{{1-\lambda_{n+1}}}  f \left(x_i\right)\]
@@
 Thus we can upper bound our previous inequality:
-
 \[f\left(\sum_{i=1}^{n+1} \lambda_i x_i\right)  \leq (1-\lambda_{n+1})\sum_{i=1}^n \frac{\lambda_i}{{1-\lambda_{n+1}}}  f \left(x_i\right) + \lambda_{n+1} f\left( x_{n+1}\right) =\sum_{i=1}^{n+1} \lambda_i f \left(x_i\right) \]
@@
 \subsection*{2.2.2}
-The characterisation of convexity states:
-
+The characterisation for differential functions for convexity states:
 \[\text{A function } f \text{ is convex} \leftrightarrow \langle \nabla f(x) - \nabla f(y), x-y \rangle \geq 0\]
@@
 For $f(x) = -\log(x)$, $\nabla f(x) = \frac{-1}{x}$:
-
 \[\langle \nabla f(x) - \nabla f(y), x-y \rangle = \langle \frac{1}{y}-\frac{1}{x}, x-y \rangle\]
 
-If $x>y$ we have the following:
+If $x>y$:
+\[x-y>0 \text{ and }\frac{1}{y} - \frac{1}{x} > 0\]
 
-\begin{gather*}
-    x-y>0\\
-    \frac{1}{y} - \frac{1}{x} > 0\\
-\end{gather*}
+so $\langle \frac{1}{y}-\frac{1}{x}, x-y \rangle > 0$.
 
-and
-\[\langle \frac{1}{y}-\frac{1}{x}, x-y \rangle > 0\]
+If $x<y$:
+\[x-y<0 \text{ and }\frac{1}{y} - \frac{1}{x} < 0\]
 
-If $x<y$ we have the following:
-\begin{gather*}
-    x-y<0\\
-    \frac{1}{y} - \frac{1}{x} < 0\\
-\end{gather*}
+so $\langle \frac{1}{y}-\frac{1}{x}, x-y \rangle > 0$.
 
-and
-\[\langle \frac{1}{y}-\frac{1}{x}, x-y \rangle > 0\]
+If $x=y$:
+\[x-y=0 \text{ and }\frac{1}{y} - \frac{1}{x} = 0\]
 
-If $x=y$ we have the following:
-\begin{gather*}
-    x-y=0\\
-    \frac{1}{y} - \frac{1}{x} = 0\\
-\end{gather*}
-
-and
-\[\langle \frac{1}{y}-\frac{1}{x}, x-y \rangle = 0\]
+so $\langle \frac{1}{y}-\frac{1}{x}, x-y \rangle = 0$.
 
@@
 Given Jensen's inequality in 2.2.1 for a convex function $f(x)$:
-
 \[f\bigg(\sum_{i=1}^n \lambda_i x_i\bigg) \leq \sum_{i=1}^n \lambda_i f(x_i)\]
@@
 and having proved in 2.2.2 that $f(x) = -\log(x)$ is convex, we can write:
-
 \[-\log\bigg(\sum_{i=1}^n \lambda_i x_i\bigg) \leq - \sum_{i=1}^n \lambda_i \log(x_i)\]
@@
 Rearranging:
-
 \[\sum_{i=1}^n \lambda_i x_i \geq \exp \bigg(\sum_{i=1}^n \log(x_i^{\lambda_i}) \bigg)\]
@@
 Choosing $\lambda_i = \frac{1}{n}$:
-
 \[\frac{1}{n}\sum_{i=1}^n x_i \geq \exp \bigg(\sum_{i=1}^n \log(x_i^{\frac{1}{n}}) \bigg)\]
 
+so $\sum_{i=1}^{n} \lambda_i = 1$ as required by Jensen's inequality.
+
 The sum of logarithms is the logarithm of the products:
-
 \[\frac{1}{n}\sum_{i=1}^n x_i \geq \exp \bigg(\log((x_1 \cdots x_n )^{\frac{1}{n}})\bigg)\]
@@
 Thus we get our inequality:
-
 \[\frac{1}{n}\sum_{i=1}^n x_i \geq \sqrt[n]{x_1 \cdots x_n}\]
@@
 Given a polytope $C=co(a_1, ..., a_m)$ in $X$, we know that any point $x \in C$ can be expressed:
-
 \[x = \sum_{i=1}^{m} \lambda_i a_i \]
@@
 where $\sum_{i=1}^{m} \lambda_i = 1$. Knowing that $f$ is a convex function on $C$, we can write Jensen's inequality:
-
 \[f\left(\sum_{i=1}^{m} \lambda_i a_i \right) \leq \sum_{i=1}^{m} \lambda_i f(a_i)\]
@@
 Because $\sum_{i=1}^{m} \lambda_i f(a_i)$ is a weighted average of $f(a_i)$'s we know that:
-
 \[\sum_{i=1}^{m} \lambda_i f(a_i) \leq \max_{i} f(a_i)\]
@@
 the weighted average will always be less than or equal to the maximum $f(a_i)$ value. Thus, we know that:
-
 \[f\left(\sum_{i=1}^{m} \lambda_i a_i \right) \leq \max_{i} f(a_i)\]
@@
 To prove that the function $f(x, y) = \| x-2y\|^2$ is convex, we will show that the Hessian is a positive semi-definite matrix. The Hessian is defined as:
-
 \[
@@
 Calculating each term:
-
 \[\nabla^2 f(x, y) = \begin{bmatrix}
@@
 
-we have our Hessian matrix for which the determinant is $(2)(8) - (-4)^2 \geq 0$ so we see that the Hessian is positive semi-definite and so the function is jointly convex.
+we have our Hessian matrix for which the eigenvalues are greater than or equal to zero, so we see that the Hessian is positive semi-definite and so the function is jointly convex.
 
@@
 Thus we have our primal problem:
-
 \[\min_{\textbf{x}}  \frac{1}{2}\|\textbf{x}\|^2 + i_{\frac{1}{\epsilon}\|\textbf{A}\textbf{x} - \textbf{b}\|_{\infty} \leq 1}\left(\textbf{Ax}\right)\]
@@
 Thus:
-
 \[g^*(\textbf{Ax}) = i^*_{\frac{1}{\epsilon}\|\textbf{A}\textbf{x} - \textbf{b}\|_{\infty} \leq 1}\left(\textbf{Ax}\right) = \frac{1}{\epsilon} \| \textbf{A}\textbf{x} - \textbf{b}\|_1\].
@@
 The Fenchel conjugate of $f(\textbf{x})$:
-
 \[f^*(\textbf{x}) =  \frac{1}{2}\|\textbf{x}\|^2\]
@@
 
-Moreover, for $g(\textbf{u}) =  i_{\frac{1}{\epsilon}\|\textbf{u} - \textbf{b}\|_{\infty} \leq 1}\left(\textbf{u}\right)$:
+Moreover, for $g(\textbf{Ax}) =  i_{\frac{1}{\epsilon}\|\textbf{Ax} - \textbf{b}\|_{\infty} \leq 1}\left(\textbf{Ax}\right)$:
 \[\textbf{0}_n \in dom(g)\]
@@
 and equivalently,
-
 \[-\textbf{A}^*\textbf{u} \in \partial f(\textbf{x}) \text{ and } \textbf{u} \in \partial g(\textbf{Ax})\]
 
-we will use the formulation involving $\partial f^*$ and $\partial g^*$, the subgradients of $f^*$ and $g^*$ respectively.
+though we will use the formulation involving $\partial f^*$ and $\partial g^*$, the subgradients of $f^*$ and $g^*$ respectively.
 
@@
 
-where $(\textbf{Ax})_i$ is the $i^{th}$ element of $\textbf{Ax}\in \mathbb{R}^{n \times 1}$
+where $(\textbf{Ax})_i$ is the $i^{th}$ element of $\textbf{Ax}\in \mathbb{R}^{n \times 1}$ for $i=1, ..., n$.
 
@@
 where $\bar{\textbf{x}} \in \mathbb{R}^{d \times 1}$ and $\bar{\textbf{u}} \in \mathbb{R}^{n \times 1}$ are minimizers for the primal and dual solution respectively, $\textbf{x}_k \in \mathbb{R}^{d \times 1}$ and $\textbf{u}_k \in \mathbb{R}^{n \times 1}$ are respectively the primal and dual solutions for the $k^{th}$ step of FISTA, and $\mu$ is the coefficient of strong convexity for $f$.
-For our problem, $\mu = 1$.
 Moreover, $\Psi(\textbf{u})=\frac{1}{2}\textbf{u}^*\textbf{A}\textbf{A}^*\textbf{u} + \frac{1}{\epsilon} \| \textbf{u} - \textbf{b}\|_1$, our dual objective.
 
-We also know for FISTA that (proof in lecture):
+We also know for FISTA that:
 \[\Psi(\textbf{u}_k) - \Psi(\bar{\textbf{u}}) \leq \frac{\|\textbf{u}_0 - \bar{\textbf{u}}\|^2}{2 \gamma \left(t_{k-1}}\right)^2\]
@@
 where $\textbf{u}_0 \in \mathbb{R}^{n \times 1}$ is the initial starting point for FISTA, and $0 < \gamma \leq \frac{1}{L}$  and $t_{k-1}$ are both scalars dictating step sizes in FISTA ($t_{k-1}$ is for step $k-1}$).
-$L$ is the Lipschitz smoothness of $f$.
-For our problem, $L=1$.
 
 Thus:
-\[ \|\textbf{x}_k - \bar{\textbf{x}}\|^2 \leq \frac{\|\textbf{u}_0 - \bar{\textbf{u}}\|^2}{4 \gamma \left(t_{k-1}}\right)^2\]
+\[ \|\textbf{x}_k - \bar{\textbf{x}}\|^2 \leq \frac{\mu \|\textbf{u}_0 - \bar{\textbf{u}}\|^2}{4 \gamma \left(t_{k-1}}\right)^2\]
 
 Simplifying:
-\[ \|\textbf{x}_k - \bar{\textbf{x}}\| \leq \frac{\|\textbf{u}_0 - \bar{\textbf{u}}\|}{2  \sqrt{\gamma}t_{k-1}}\]
+\[ \|\textbf{x}_k - \bar{\textbf{x}}\| \leq \sqrt{\frac{\mu}{4\gamma}}\frac{\|\textbf{u}_0 - \bar{\textbf{u}}\|}{t_{k-1}}\]
 
@@
 
-where $t_0  =1 $
-
 In both cases the rate of convergence of the primal iterates when applying FISTA on the dual problem:
-\[ \|\textbf{x}_k - \bar{\textbf{x}}\| \leq \frac{a\|\textbf{u}_0 - \bar{\textbf{u}}\|}{\sqrt{\gamma}(k+1)}\]
+\[ \|\textbf{x}_k - \bar{\textbf{x}}\| \leq \sqrt{\frac{a^2 \mu}{4\gamma}}\frac{\|\textbf{u}_0 - \bar{\textbf{u}}\|}{k+1}\]
 
@@
 The Proximal Stochastic Gradient Algorithm:
-
 \[\textbf{x}^{k+1} = prox_{\gamma_k \lambda \|\cdot\|_1} \left(\textbf{x}^k - \gamma_k \left(\left\langle \textbf{a}^{i_k}, \textbf{x}^{k}\right\rangle - y_{i_k}\right)\textbf{a}^{i_k}\right)\]
@@
 where:
-
 \[\gamma_k = \frac{n}{\|A\|^2 \sqrt{k+1}}\]
@@
   \centering
-\includegraphics[scale=0.3]{outputs/part_3/psga-loss}
+\includegraphics[scale=0.4]{outputs/part_3/psga-loss}
 \caption{PSGA Loss}
@@
   \centering
-\includegraphics[scale=0.3]{outputs/part_3/rcpga-loss}
+\includegraphics[scale=0.4]{outputs/part_3/rcpga-loss}
 \caption{RCPGA Loss}
@@
 The Fast Iterative Shrinkage Threshold Algorithm (FISTA):
-
 \[\alpha_{k+1} = prox_{\gamma g^*(\cdot)}\bigg(\nu_k + \gamma \textbf{A}_y \nabla f^*(-\textbf{A}_y^* \alpha_{k})\bigg)\]
-
 \[\nu_{k+1} = \alpha_{k} + \frac{t_k - 1}{t_{k+1}}(\alpha_{k+1}-\alpha_{k})\]
 
-where $t_k = \frac{1+\sqrt{1+ 4 (t_{k-1})^2}}{2}$ or $t_k = \frac{k+a}{a}$ with $a \geq 2$. For our implementation we chose the first definition for $t_k$.
+where we choose $t_k = \frac{1+\sqrt{1+ 4 (t_{k-1})^2}}{2}$ for our implementation.
 
 For our dual problem:
-
-\[\nabla f^*(-\textbf{A}_y^* \alpha_{k})) = -\textbf{A}_y^* \alpha_{k}\]
+\[\nabla f^*(-\textbf{A}_y^* \alpha_{k}) = -\textbf{A}_y^* \alpha_{k}\]
 
 and
-
 \[prox_{\gamma g^*(\cdot)}(\omega) = \begin{cases}
@@
 
-With FISTA implemented, we can plot the dual objective function:
-
+Implementing FISTA, we can plot the dual objective function:
 \begin{figure}[h]
 \centering
-\includegraphics[scale=0.3]{outputs/part_4/fista-loss}
+\includegraphics[scale=0.35]{outputs/part_4/fista-loss}
 \caption{FISTA Loss}
@@
 \end{figure}
-
 \subsection*{4.2}
-
 We can also implement the Randomised Coordinate Projected Gradient Algorithm on the dual problem and plot the dual objective function:
-
 \begin{figure}[h]
 \centering
-\includegraphics[scale=0.3]{outputs/part_4/rcpga-loss}
+\includegraphics[scale=0.35]{outputs/part_4/rcpga-loss}
 \caption{RCPGA Loss}
@@
 We can plot the decision boundary for the randomly initialised $\alpha$ vector as well as the two classes:
-
 \begin{figure}[h]
 \centering
-\includegraphics[scale=0.3]{outputs/part_4/initial-contour}
+\includegraphics[scale=0.28]{outputs/part_4/initial-contour}
 \caption{Initial Contour Plot}
@@
 Similarly for each algorithm, we can plot the decision boundaries from the learned $\alpha$:
-
 \begin{figure}[h]
@@
   \centering
-\includegraphics[scale=0.3]{outputs/part_4/rcpga-contour}
+\includegraphics[scale=0.28]{outputs/part_4/rcpga-contour}
 \caption{RCPGA Contour Plot}
@@
   \centering
-\includegraphics[scale=0.3]{outputs/part_4/fista-contour}
+\includegraphics[scale=0.28]{outputs/part_4/fista-contour}
 \caption{FISTA Contour Plot}
@@
 Although each FISTA step involves two steps (to calculate $\alpha_{k+1}$ and $\nu_{k+1}$), this is still a much faster algorithm than RCGPA.
-However, we can also see that the objective function for FISTA is much less smooth compared to RCGPA.
+However, we can also see that the objective function for FISTA with respect to the algorithm iteration is much less smooth compared to RCGPA's loss curve.
 Depending on use case, this may or may not be a problem.
 Overall, FISTA is able to produce quality decision boundaries much faster than RCGPA.
+Moreover, during the parameter turning process of $\lambda$ and $\sigma$, RCPGA was much more robust, converging to good decision boundaries most of the time when compared to FISTA, which required more tuning of parameters to converge to a decision boundary that perfectly separated the two classes of the training data.
 

### 2023-01-08 src/problems.py

@@ -164,9 +164,12 @@ class SparseProblem(Problem):
         :param lambda_parameter: scalar on g(x) of the minimization problem
         :return: LASSO loss evaluation at x with lambda
         """
-        return 0.5 * np.mean(
-            (self.a_matrix @ x - self.y) ** 2
-        ) + lambda_parameter * np.linalg.norm(x, ord=1)
+        return (
+            (1 / (2 * self.n))
+            * (self.a_matrix @ x - self.y).T
+            @ (self.a_matrix @ x - self.y)
+            + lambda_parameter * np.linalg.norm(x, ord=1)
+        ).item()
 
     def grad_f_j(self, x: np.ndarray, j: int) -> float:
         """

### 2023-01-08 part_4.py
update contour plots

@@ -131,7 +131,7 @@ def part_4(
 
 if __name__ == "__main__":
     np.random.seed(DEFAULT_SEED)
-    CONTOUR_RESOLUTION = 250
+    CONTOUR_RESOLUTION = 1000
     part_4(
         contour_resolution=CONTOUR_RESOLUTION,
     )

### 2023-01-08 submission.tex
fix typo

@@
 Overall, FISTA is able to produce quality decision boundaries much faster than RCGPA.
-Moreover, during the parameter turning process of $\lambda$ and $\sigma$, RCPGA was much more robust, converging to good decision boundaries most of the time when compared to FISTA, which required more tuning of parameters to converge to a decision boundary that perfectly separated the two classes of the training data.
+Moreover, during the parameter tuning process of $\lambda$ and $\sigma$, RCPGA was much more robust, converging to good decision boundaries most of the time when compared to FISTA, which required more tuning of parameters to converge to a decision boundary that perfectly separated the two classes of the training data.
 

### 2023-01-09 submission.tex
minor report edits

@@
 \[ f(x) = \begin{cases}
-      - (1 + \log (-u), & \text{if } u < 0 \\
+      - (1 + \log (-u)), & \text{if } u < 0 \\
       +\infty, & \text{if } u \geq 0 \\
@@
       -\frac{1}{\epsilon}, &  \text{if } | u_i - b_i| < 0 \\
-      [-\frac{1}{\epsilon}, \frac{1}{\epsilon}], &  \text{if } | u_i - b| = 0 \\
+      [-\frac{1}{\epsilon}, \frac{1}{\epsilon}], &  \text{if } | u_i - b_i| = 0 \\
    \end{cases} \\
@@
       -\frac{1}{\epsilon}, &  \text{if } | u_i - b_i| < 0 \\
-      [-\frac{1}{\epsilon}, \frac{1}{\epsilon}], &  \text{if } | u_i - b| = 0 \\
+      [-\frac{1}{\epsilon}, \frac{1}{\epsilon}], &  \text{if } | u_i - b_i| = 0 \\
    \end{cases} \\\]
@@
 It is difficult to determine which of the two decision boundaries are more suitable for the training data, without more context about the data itself.
-However considering the loss function plots for both algorithms, we can see that FISTA is able to converge in around 20 iterations whereas RCPGA required on the order of $3\times 10^{5}$ iterations.
+However considering the loss function plots for both algorithms, we can see that FISTA was able to converge in around 20 iterations whereas RCPGA required on the order of $3\times 10^{5}$ iterations.
 Although each FISTA step involves two steps (to calculate $\alpha_{k+1}$ and $\nu_{k+1}$), this is still a much faster algorithm than RCGPA.
 However, we can also see that the objective function for FISTA with respect to the algorithm iteration is much less smooth compared to RCGPA's loss curve.
-Depending on use case, this may or may not be a problem.
+Depending on use case, this may or may not be less desirable.
 Overall, FISTA is able to produce quality decision boundaries much faster than RCGPA.
-Moreover, during the parameter tuning process of $\lambda$ and $\sigma$, RCPGA was much more robust, converging to good decision boundaries most of the time when compared to FISTA, which required more tuning of parameters to converge to a decision boundary that perfectly separated the two classes of the training data.
+However, during the parameter tuning process of $\lambda$ and $\sigma$, RCPGA was much more robust, converging to good decision boundaries most of the time when compared to FISTA, which required more tuning of parameters to converge to a decision boundary that perfectly separated the two classes of the training data.
 
